### 2015-08-11 paper/manuscript.tex
Add paper and add top-level directory

\documentclass[11pt]{amsart}
\usepackage{graphicx,amssymb,amsthm,setspace,lineno}
\usepackage[round]{natbib}
\linenumbers
\onehalfspacing
\title{Estimating Admixture Parameters from Admixture Fractions}
\author{Mason Liang and Rasmus Nielsen}
\date{\today}

\begin{document}
\begin{abstract}
Estimation of admixture fractions has become one of the most commonly used computational tools in population genomics.  However, there is remarkably little population genetic theory on their statistical properties.  We develop theoretical results that can accurately predict means and variances of admixture proportions within a population using models with recombination and genetic drift. Based on established theory on measures of multilocus disequilibrium, we show that there is a set of recurrence relations that can be used to derive expectations for higher moments of the admixture fraction distribution. We obtain closed form solutions for some special cases. Using these results, we develop a method for estimating admixture parameters from estimated admixture proportion obtained from programs such as Structure or Admixture. We apply this method to HapMap data and find that the population history of African Americans, as expected, is not best explained by a single admixture event between people of European and African ancestry. A model of constant gene flow for the past 11 generations until 2 generations ago gives a better fit.
\end{abstract}
\maketitle

\section*{Introduction}
It is common in population genetic analyses to consider individuals as belonging fractionally to two or more discrete source populations. The proportion of an individual's genome that belongs to a population is called that individual's ``admixture fraction" or ``admixture proportion". Programs such as \verb|Structure| \citep{pritchard2000inference}, \verb|Eigenstrat|, \verb|Frappe| \citep{tang2005estimation}, or \verb|Admixture| \citep{alexander2009fast} can jointly estimate these admixture fractions for multiple individuals in a sample, along with the corresponding  allele frequencies in each of the source populations. These admixture fractions are often presented in a `structure plot,' an example of which is shown in Figure \ref{structure_plot}. We will henceforth refer to these methods as 'structure analyses'. 
This approach has proven highly useful for understanding genetic relationships in many different species, e.g. humans \citep{rosenberg2002genetic}, cats \citep{menotti2008patterns}, or pandas \citep{zhang2007genetic}. Other analyses reconstruct admixture tracts for each genome in the sample, by inferring the local ancestry of every position, or window, in each sampled genome \citep{tang2006reconstructing,maples2013rfmix}. In this context, the admixture fraction for a genome is the fraction of its total length that is inherited from a particular source population.

Although structure analyses are not tied to any particular mechanistic model of population history and demography, the admixture fractions and admixture tracts are commonly interpreted to be the result of past admixture events in which modern populations were formed by admixture (or introgression) between ancestral source populations. The distribution of admixture tract lengths has been related to specific mechanistic models of admixture \citep{falush2003inference,tang2006reconstructing,pool2009inference}, and has been used to estimate times of admixture \citep{gravel2012population}. However, the admixture proportions themselves also contain information regarding admixture times. Right after an admixture event, the variance in admixture proportions within a population will be high, but will thereafter decrease, and will eventually (as we will show) converge to zero in the limit of large genomes.  
The variance in admixture fractions among individuals contains substantial information about the time since admixture that can be used in addition to the tract length distribution.  
It may in some cases be more robust than inferences based on tract lengths, because the length distribution of tracts is often difficult to infer, and is often not modeled accurately by the hidden Markov model (HMM) methods used to infer tract lengths \citep{liang2014lengths}.
Even in cases where tract lengths can be accurately inferred, studies aimed at estimating admixture times should benefit from using both variance in admixture proportions among individuals and overall admixture tract lengths distributions.

\citet{verdu2011general} developed a method for computing moments of admixture proportions in a model in which admixed population is formed as a mixture between multiple source populations, allowing for arbitrary gene-flow from the source populations over a number of generations ($g$). They establish recursions for the moments of the admixture fractions and use these equations to determine how the mean and the variance changes through time in particular admixture scenarios. These moments are expectations for \textit{single} individual's admixture fraction and are averaged over the possibile genealogical histories of the population. As a result, they can be difficult to relate to data because replicates from multiple identical populations rarely are available. In this paper, we consider a different problem, the problem of calculating sample moments for admixture proportions obtained from individuals in one population.

We extend the model model in \citet{verdu2011general} to incorporate the effects of recombination and genetic drift by adding a a random union of gametes component. Recombination is important because even if one half of a chromosome's ancestors are from the first source population, it is unlikely that exactly one half of that chromosome's genetic material is inherited from that population. Genetic drift is important because the individuals in a sample might share ancestors and, therefore, have more similar admixture fractions than expected by chance in a model without drift. The results developed in this paper should be directly applicable for quantifying the results of a structure analysis.

\section*{The General Mechanistic Model}
We start by considering admixture fractions in haploid genomes. These haploid admixture fractions can later be paired up to create diploid admixture fractions. The admixture fraction of a (haploid) genome $H_i$, is the proportion of $H_i$ that is inherited from a particular source population. For notational simplicity, we only consider gene-flow only from one population into another. We will later discuss how to extend this model to multiple admixing source populations. We use the same mechanistic admixture model of \citet{verdu2011general}, and will use its notation where possible. Finally, we use the random union of gametes model for genetic drift and recombination, and assume a sample size of $n$ chromosomes from a single population.

In the model of \citet{verdu2011general}, a hybrid population of $N$ diploid individuals forms in generation 1 from two isolated source populations. In this first generation, individuals in the hybrid population are from the first source population with probability $s_0$ or from the second source population with probability $1-s_0$. In subsequent generations, each individual in the hybrid population draws two recombined gametes at random from the previous generation. In generation $g+1$, each gamete is, independently, from the first source population with introgression probability $s_g$, or from the hybrid population with probability $1-s_g$. If the individual inherits both of its gametes from the hybrid population, the two gametes are from the same individual with probability $1/N$, and from different individuals with probability $1-1/N$.

Mathematically, if the stochastic process $A(\ell)$ represents the local ancestry along a chromosome as a function of physical position ($\ell$):

$$
A(\ell) = \left\{
     \begin{array}{l}
       0\ :\ \ell $ descended from first source population$\\
       1\ :\ \ell $ descended from second source population$\\
     \end{array}
   \right.,
$$

then the fraction of the chromosome descended from the second source population is given by

$$
H=\frac{1}{L}\int_0^LA(\ell) d\ell,
$$

where $L$ is the total length of the chromosome.

Assume that $g$ generations after the start of admixture we have randomly sampled $n$ chromosomes from the hybrid population and determined their corresponding admixture fractions, $H_{1(g)},H_{2(g)},\dots,H_{n(g)}$.
We are interested in the joint distribution of these $n$ random variables.
When $n=1$ and as $L\rightarrow\infty$, this is the admixture fraction considered in \citep{verdu2011general}.

Because the $n$ chromosomes have possibly overlapping geneologies, the admixture fractions are not independent.
However, the joint distribution of the admixture fractions does not depend on their ordering, so they are exchangeable (see appendix for mathematical definition). 
As a result, they can be viewed as being identically and independently ($\textit{iid}$) drawn from a random distribution $\mathcal{G}$.
This random distribution can be interpreted as a deterministic function of the random genealogy of the entire hybrid population up to $g$ generations in the past.
When $g$ is small, the genealogies of the $n$ samples will be unlikely to differ from $n$ non-overlapping binary trees, so $\mathcal{M}$ will be approximately constant. If $g$ is large however, this will no longer hold.

\citet{verdu2011general} focus on moments of $H_{1(g)}$, in particular on the mean and variance.
However, because the admixture fractions are not independent, even as $n\rightarrow\infty$, the sample mean and sample variance will converge to the mean and variance of $\mathcal{G}$, which are random quantities.
For example,

\begin{align*}
\mathbb{E}(H_{1(g)})&\neq\mathbb{E}(H_{1(g)}|\mathcal{M})=\lim_{n\rightarrow\infty}\frac{1}{n}\sum_{i=1}^n H_{i(g)}\\
\text{var}\left(H_{1(g)}\right)&\neq\text{var}(H_{1(g)}|\mathcal{M})=\lim_{n\rightarrow\infty}\frac{1}{n-1}\sum_{i=1}^n \left(H_{i(g)}-\frac{1}{n}\sum_{j=0}^n H_{j(g)}\right)^2.
\end{align*}

and similarly for higher-order moments.  In other words, the moments of the admixture factions have two components: randomness due sampling of population history, and randomness due to the sampling of individuals.  The expressions to left account for both, while the expressions to the right only account for the latter.  Variances among individuals within a population will correspond to var$(H_{1(g)}|\mathcal{M})$ but var$(H_{1(g)})$ will tend to be larger than the expected sample variance calculated from multiple individuals sampled from the same population, and will rarely be useful for inference purposes.

In the following sections, we will show how the constants on the left-hand side, as well as expectations of the random variables on the right-hand side, can be derived for mechanistic models of introgression.
By comparing these expectations to the observed admixture parameters from a sample, we will be able to construct a method of moments estimator for the  parameters of the model.

We let $k_1$ be the sample mean:

$$
	k_1\equiv \frac{1}{n}\sum_{i=1}^n H_{i(g)}.
$$

We can express its expectation in terms of the 1-point correlation function of $A$:

\begin{align*}
	\mathbb{E}(k_1)
	&=\mathbb{E}(H_{1(g)})\\
	&=\frac{1}{L}\int_0^L \mathbb{P}\{A_{1(g)}(\ell)=1\}d\ell\\
	&=\mathbb{P}\{A_{1(g)}(0)=1\}.
\end{align*}

We let $k_2$ be the unbiased estimator of the sample variance:

\begin{align*}
k_2&\equiv\frac{1}{n-1}\sum_{i=1}^n\left(H_{i(g)}-k_1\right)^2.
\end{align*}

Its expectation is given by

\begin{align*}
	\mathbb{E}(k_2)
	&=\frac{1}{n-1}\sum_{i=1}^n \mathbb{E}(H_{i,g}^2)-\frac{1}{n(n-1)}\sum_{i,j=1}^n \mathbb{E}(H_{i,g}H_{j,g})\\
	&=\mathbb{E}(H_{1,g}^2)-\mathbb{E}(H_{1,g}H_{2,g}).
\end{align*}

These expectations can be written in terms of two-point correlation functions:

\begin{align*}
	\mathbb{E}(H_{1(g)}^2)
		&=\frac{1}{L^2}\mathbb{E}\left(\int_0^L A_{1(g)}(\ell) d\ell \int_0^L A_{1(g)}(\ell) d\ell\right)\\
		&=\frac{1}{L^2}\int_0^L\int_0^L\mathbb{E}\left(A_{1(g)}(\ell) A_{1(g)}(\ell')\right) d\ell d\ell'\\
		&=\frac{1}{L^2}\int_0^L\int_0^L\mathbb{P}\left\{A_{1(g)}(\ell)=1, A_{1(g)}(\ell')=1\right\} d\ell d\ell'.
\end{align*}

Similarly,

\begin{align*}
	\mathbb{E}(H_{1(g)}H_{2(g)})
		&=\frac{1}{L^2}\int_0^L\int_0^L\mathbb{P}\left\{ A_{1(g)}(\ell)=1, A_{2(g)}(\ell')=1\right\} d\ell d\ell'.
\end{align*}

Writing these two correlation functions as

$$\mathbf{v}_{2(g)}=\left( \begin{array}{r}
			\mathbb{P}\left\{A_{1,g}(\ell)=1, A_{1,g}(\ell')=1\right\} \\ \mathbb{P}\left\{A_{1,g}(\ell)=1, A_{2,g}(\ell')=1\right\}
		\end{array} \right),$$

we get that

\begin{align}
	\mathbb{E}(k_2)&=\frac{1}{L^2}\int_0^L\int_0^L\left( \begin{array}{cc}
			1 & -1	\end{array} \right)\mathbf{v}_{2(g)} d\ell d\ell'.
	\label{k2}
\end{align}

In general, the $i^\text{th}$ $k$-statistic is an unbiased estimator of the $i^\text{th}$ cumulant of $\mathcal{M}$, and its expectation can be written as an integral over $[0,L]^i $ of a linear combinations of $i$-point correlation functions.
For example,

\begin{align*}
	\mathbb{E}(k_3)&=\frac{1}{L^3}\int_0^L\int_0^L\int_0^L\left( \begin{array}{ccccc}
			1 & -1 &-1 &-1 &2	\end{array} \right)\mathbf{v}_{3(g)} d\ell d\ell'd\ell''\\
	\mathbb{E}(k_4)&=\frac{1}{L^4}\int_0^L\int_0^L\int_0^L\int_0^L \Bigl(
			1 \  \underbrace{-1}_\text{4 times} \  \underbrace{-1}_\text{3 times} \  \underbrace{2}_\text{6 times} \ 6 \Bigr)\mathbf{v}_{4(g)} d\ell d\ell'd\ell''d\ell'''\\
			\dots
\end{align*}

Remarkably, the linear combinations required to compute the expectations of the $k$-statistics correspond exactly to the higher-order disequilibria as defined in [Bennett 1948?].
Furthermore, if instead the we choose to compute the expectations of the $h$-statistics, which estimate the central moments, the linear combinations would correspond to the higher-order disequilibria as defined in [Slatkin 1972].

We next find the recurrence relations these correlation functions satisfy and solve them in the some special cases. In particular we will consider the case of a single admixture event $g$ generations ago and the case of constant gene-flow starting $g$ generations ago.

\subsection*{A Single Admixture Event}
We start with a simple case, where introgression only occurs in the founding generation, i.e. $s_g=0$ for $g>0$. Using the random union of gametes model, we can compute $\mathbf{v}_{2(g)}$ in terms of the probabilities from the previous generation:

If two sites at $\ell$ and $\ell'$ are on the same chromosome in generation $g$, then they were inherited from one gamete from one parent in the previous generation, while if they 	are on different haplotypes, then the probability that they are from gametes from one parent is $1/N$.
On one gamete, the probability that these two sites are inherited from the same parental chromosome is $[\ell \ell']$ and the probability that they are instead inherited from different parental chromosomes is $[\ell|\ell']$.
In matrix notation,

\begin{align*}
\mathbf{v}_{2(g)}&=
		\left(\mathbf{L}_2\mathbf{U}_2\right)^g
		\mathbf{v}_{2(0)},
\end{align*}

where the the recombination and drift matrices are given by

\begin{align*}
	\mathbf{U}_2&=\left( \begin{array}{cc}
					 {[\ell \ell']} & [\ell|\ell']\\
					 0 & 1
				\end{array} \right)\\
	\mathbf{L}_2&=\left( \begin{array}{cc}
					 1 & 0\\
					 1/N & 1-1/N
				\end{array} \right).
\end{align*}

This is the the same matrix equation (Wright 1933 and Hill and Robertson 1966) derived for the decay of two-locus linkage disequilibrium.
The ``alleles" we consider are the local ancestry at $\ell$ and $\ell'$.
To the extent possible, our notation will follow (Hill 1974), whose results for measures of multi-locus linkage disequilibria we use.
The matrices $\mathbf{L}_2$ and $\mathbf{U}_2$ share $(1\ -1)$ as a left-eigenvector, with corresponding eigenvalues $1-1/N$ and $[\ell\ell']$.
As a result,

\begin{align}
	\mathbb{E}(k_2)&=\frac{1}{L^2}\int_0^L\int_0^L\left( \begin{array}{cc}
			1 & -1	\end{array} \right)\cdot\left(\mathbf{L}_2\mathbf{U}_2\right)^g
		\mathbf{v}_{2(0)} d\ell d\ell'\nonumber\\
		&=\frac{1}{L^2}\left(1-\frac{1}{N}\right)^g\left( s_{1,0}-s_{1,0}^2 \right)\int_0^L\int_0^L [\ell\ell']^g d\ell d\ell'.
	\label{k2_pulse}
\end{align}

For a model using the Haldane map function,  $[\ell|\ell']=\frac{1-\exp(-2|\ell-\ell'|)}{2}$, this equation becomes

\begin{align*}
	\mathbb{E}(k_2)&=\frac{1}{L^2}\left(1-\frac{1}{N}\right)^g\left( s_{1,0}-s_{1,0}^2 \right)\int_0^L\int_0^L \left(\frac{1+\exp(-2|\ell-\ell'|)}{2}\right)^g d\ell d\ell',
\end{align*}

 while for a model of complete crossover inteference on a chromosome of length 1 Morgan, we can get a closed form solution:

\begin{align*}
	\mathbb{E}(k_2)&=\left(1-\frac{1}{N}\right)^g\left( s_{1,0}-s_{1,0}^2 \right)\int_0^1\int_0^1 \left(1-|\ell-\ell'|\right)^g d\ell d\ell'\\
	&=\left(1-\frac{1}{N}\right)^g \left(s_{1,0}-s_{1,0}^2\right) \frac{2}{2+g}.
\end{align*}

For predicting the expected sample variance, the difference between these two models is not large, as shown in figure \ref{ek2_rec}.

For the simulations and inference in this paper, we will ignore crossover interference, and use the Haldane map function.
However, none of the mathematical results of this paper will require this assumption.

For computing higher-order correlation functions, we find a similar equation

\begin{align}
\mathbf{v}_{i(g)}&=
		\left(\mathbf{L}_i\mathbf{U}_i\right)^g
		\mathbf{v}_{i(0)}.
\label{rug}
\end{align}

Bennett's coefficients for higher-order linkage are left-eigenvectors of the recombination matrix $\mathbf{U}_i$. 
For $i=3$, it is also a left-eigenvector of the drift matrix, so we immediately get that

\begin{align*}
	\mathbb{E}(k_3)&=\frac{s_0(1-s_0)(2-s_0)}{L^3}\left(1-\frac{1}{2N}\right)^T\left(1-\frac{2}{2N}\right)^T\int_0^L\int_0^L\int_0^L[\ell\ell'\ell'']^Td\ell d\ell'd\ell''.
\end{align*}

For $i\geq 4$, this is no longer true, but the results of [Hill 197?] can be used to compute $\textbf{v}_{i}$ without having to exponentiate the entire drift and recombination matrices. For example, for $k_4$, the drift and recombination matrices are $15\times 15$, but using the technique in [Hill 197?], we only need to exponentiate a $4\times 4$ matrix to compute $\mathbb{E}(k_4)$.
%[Comment: I didn't understand the end of this sentence].

\subsection*{Varying Migration}
If $s_g>0$ for $s\geq 1$, we obtain a modified version of Equation \ref{rug}:

\begin{align}
\mathbf{v}_{i(g)}&=
		\mathbf{L}_i \mathbf{D}_{i(g)}\mathbf{U}_i
		\mathbf{v}_{i(g-1)},
		\label{with_migration}
\end{align}

where the diagonal matrix $\mathbf{D}_{i(g)}$ has entries giving the probabilities that all $|p|$ haplotypes in a correlation function are not inherited from migrant ancestors in the previous generation:

$$ d_{p,p(g)} = (1-s_g)^{|p|}.$$

Note that as $s_{(g)}$ is a constant, then equation (\ref{with_migration}) is linear, and can be solved using a Laplace transform.

%[insert stuff here]

\section*{Inference of admixture times}
The equations in the previous section can be used to develop a method of moments-estimators for admixture parameters by numerically solving the admixture parameters in terms of the expectations for the $k$-statistics. Substituting in the observed values for the $k$-statistics gives estimates for the admixture parameter(s).

However, with real data, we only have estimates of the admixture fractions, so some of the variability seen in the distribution of admixture fractions will be due to estimation variability.
To account for this, we assume that the estimations errors are additive and \textit{iid}:

$$\hat{H}_{i(g)}=H_{i(g)}+\epsilon_i.$$

Because cumulants are additive,

\begin{align*}
	\mathbb{E}(k_n)&=\mathbb{E}\left(\kappa_n(H_{i(g)}+\epsilon_i|\mathcal{M})\right)\\
	&=\mathbb{E}\left(\kappa_n(H_{i(g)}|\mathcal{M})\right)+\kappa_n(\epsilon_i).
\end{align*}
The expectations we have computed are just the term of this sum.
To correct for the variability in the estimates, we need to subtract off the second term. 
We use a block bootstrap to estimate these effects.

One additional complication arises in dealing with genotyping data. We have assumed that we have the ancestry fractions for each haplotype in the sample, but with genotyping data, we instead have their pairwise means: $(H_{1(g)}+H_{2(g)})/2\dots$. In turns out this is results in a decrease in the expectations of the $k$-statitics. Conditional on the random distribution $\mathcal{G}$, $H_{1(g)},H_{2(g)},\dots$ are \textit{iid} drawn from $\mathcal{G}$.
Cumulants are additive, so we use the law of total expectation to find that

\begin{align*}
	\kappa_i\left(\frac{H_{1(g)}+H_{2(g)}}{2}\right)
		&=\mathbb{E}\left(\kappa_i\left(\frac{H_{1(g)}+H_{2(g)}}{2}\middle\vert\mathcal{G}\right)\right)\\
		&=\mathbb{E}\left(\kappa_i\left(\frac{H_{1(g)}}{2}\middle\vert\mathcal{G}\right)+\kappa_i\left(\frac{H_{2(g)}}{2}\middle\vert\mathcal{M}\right)\right)\\
		&=2^{-i+1}\mathbb{E}\left(\kappa_i\left(H_{1(g)}\middle\vert\mathcal{G}\right)\right)\\
		&=2^{-i+1}\kappa_i\left(H_{1(g)}\right).
\end{align*}

\subsection*{Comparison to Verdu and Rosenberg}
The recursion equations given by \citet{verdu2011general} are different from the ones we have derived. This is partly because we have accounted for the effects of genetic drift and recombination, but also because we are computing the moments of slightly different quantities.

In figure \ref{ensemble}, we have shown the admixture fractions for five replicate populations 5, 50, and 500 generations after an admixture pulse. The variance that \citep{verdu2011general} compute variance over all the replicate populations, while the variance we have computed in this paper is the expectation of the variance within a single population. When $g$ is small, these are not very different, but when $g$ is large, the variance within a population goes to zero, but the variance across the replicate populations does not. This effect is shown in Figure \ref{ek2_verdu}. Initially, both quantities decline exponentially in $g$, but after $2^g > nLg$ 
%[comment: this I don't understand]
, the variance we predict begins to decline linearly instead. This is because variance is inversely proportional to the number of genetic ancestors of the sample. When $g$ is small, the number of genetic ancestors is approximately $2^g$. However, the approximate number of recombination events in the sample is approximately bounded by $nLg$, so when this quantity is smaller than $2^g$, it provides a better approximation for the number of genetic ancestors. In this regime, the variance will decline linearly in $g$.

It is also possible to compute the variance over all population replicates under our model, which allows a direct comparison to \citet{verdu2011general}.
In the case of one pulse of admixture, we can now solve equations \ref{k2} for $\mathbb{P}\left\{A_{1,g}(\ell)=1, A_{1,g}(\ell')=1\right\}$ to get

\begin{align}\nonumber
	\text{var}(H_{1(g)})&=\mathbb{E}(H_{1,g}^2)-s_{1,0}^2\\
		&=\frac{1}{L^2}\int_0^L\int_0^L\mathbb{P}\left\{A_{1,g}(\ell)=1, A_{1,g}(\ell')=1\right\} d\ell d\ell'-s_{1,0}^2\nonumber\\
		&=\frac{1}{L^2}\left(s_{1,0}-s_{1,0}^2\right)\int_0^L\int_0^L1-\left(1-[\ell\ell']\right)\frac{1-[\ell\ell']^g\left(1-\frac{1}{N}\right)^g}{1-[\ell\ell']\left(1-\frac{1}{N}\right)} d\ell d\ell.
		\label{var}
\end{align}

This variance and the expectation of the second $k$-statistic have the same limit as $N\rightarrow\infty$, but for finite $N$, the variance will always be larger.
This is because

\begin{align*}
\text{var}(H_{1(g)})&=
	\text{var}\left[\mathbb{E}(H_{1(g)}|\mathcal{G})\right]
	+\mathbb{E}\left[\text{var}(H_{1(g)}|\mathcal{G})\right]=\text{var}[k_1]+\mathbb{E}[k_2].
\end{align*}

The first variance is small when $N$ is large, but is always non-negative.  This function is also plotted on figure \ref{ek2_verdu}. The difference between this equation and equation \ref{k2} only becomes significant on a coalescent time scale. In the absence of genetic drift, the admixture fractions are approximately independent, becuase the samples do not share ancestors.

\subsection*{Application African American Date}
We applied this method to a subset of the ASW, CEU, and YRI data from the HapMap 3 project [citation here].
After excluding children from trios, there were the genotypes for 49 ASW, 113 YRI, and 112 CEU individuals. 
We estimated the admixture fractions using the supervised learning mode of \verb|Admixture| [citation here], with the CEU and YRI individuals assigned to separate clusters. 
The sampling distribution of the admixture fractions was estimated using the block bootstrap with $10^4$ replicates and 2678 blocks, giving a block size of approximately 10 CM.
The admixture fractions for the 49 ASW samples are shown in Figure \ref{structure_plot} and the observed $k$-statistics are given in table \ref{table:k_stats}.

We assumed a 3-parameter model of constant admixture. For $g_{start}\leq g\leq g_{stop}$, $s_g=s$ with $s_g=0$ elsewhere. By matching the block-bootstrap corrected $k_2$ and $k_3$ to the predictions of equation \ref{k2}, we obtained a point estimates of

\begin{align*}
 \hat{s} &= 0.0277\\
 \hat{g}_{start} &= 2 \\
 \hat{g}_{stop} &= 11.
\end{align*}

We obtained confidence intervals, shown in Figure \ref{ci}, by simulation. For each cell in the grid, we simulated $10^3$ replicates under the corresponding $g_{start}$ and $g_{stop}$, with $s=1-k_1^{1/(g_{stop}-g_{start}+1)}$. For each replicate, we computed the $k_2$, $k_3$, and $k_4$ statistics. A cell was then included in the confidence interval if and only if the corrected $k_2$, $k_3$, and $k_4$ statistics from the HapMap data fall inside a centered interval containing 98.7\% of the probability mass of the simulated distribution. This mass was chosen so that under the Bonferroni correction for three tests, there is at least a 95\% chance of including the true parameter values in the confidence region.

%[comment: this I don't understand.  In the figure you write 95\% - here you say 99\%?.  Also, you have three statistics, do you require each of them to fall within? I don't think that is valid.  You need to reduce to a single statistic or consider their joint distribution somehow.]

The point estimates for $g_{start}$ and $g_{stop}$ correspond to the values for which the observed $k$-statistics are closest to their simulated medians.

\section*{Discussion}

We have extended the mechanistic model of \citet{verdu2011general} to account for recombination and genetic drift. Doing so allows us to apply the predictions of this model to data. This mechanistic model allows for a large number of parameters. For the purposes of inference, it seems that imposing constraints, i.e. a small number of pulses or constant admixture, will be needed to narrow the search space.

In this paper, we have assumed that admixture only comes from one source population, this need not be the case. To account for admixture from multple source populations, equation \ref{k2} must be modified to account for the probability that haplotypes trace their descent to multiple source populations. Alogrithmically, this is feasible, but the notation is cumbersome. The resulting equations are given in the appendix, along with the equations for computing expectations of higher-order $k$-statistics.

Applications of the method to African-American HapMap data provides estimates of the time since admixture between people of Europe and and African descent in America.  Notice that the confidence set for the admixture parameters does not include values of $g_stop = 0$.  We interpret this as evidence that admixture rates have declined the last two generations.  Also notice that admixture before 15 generations ago can be rejected.  With a generation time of 25-30 years, this corresponds to 325-400 years, and is in good accordance with the historical record.  The point estimate of the time of first admixture is 11 generations, or approx. 275-330 years ago.

Structure analyses have become one of the most commonly applied tools in population genomic analyses.  The theory developed in this paper allows users of structure analyses to interpret their data in the context of an model of admixture between populations, and should find use in many studies aimed at understanding the history of populations.

%\section*{Constant Admixture}
%Suppose instead that every generation, a fraction, $m_1$, of the population is replaced by migrants from source population 1. 
%There are now three parameters, $\mathbf{\Theta}=\{T,m_0,m_1\}$.
%The recursive formulas become slightly more complex, but still solvable.
%\begin{align*}
%	\mathbb{P}\{A_{1,T}(\ell)=1\}&= m_1+(1-m_1)\mathbb{P}\{A_{1,T-1}(\ell)=1\}\\
%		&=1-(1-m_1)^T\left(1-\mathbb{P}\{A_{1,T}(\ell)=1\}\right)\\
%		&=1-(1-m_1)^T\left(1-m_0\right).
%\end{align*}
%\begin{align*}
%	\mathbf{P}_T&=
%		\mathbf{L}_2\mathbf{U}_2
%		\left(\begin{array}{cc}1-m_1 & 0 \\0 & (1-m_1)^2\end{array}\right)
%		\mathbf{P}_{T-1}+\left(\begin{array}{c}m_1 \\ m_1^2+2m_1\mathbb{P}\{A_{1,T-1}(\ell)=1\}\end{array}\right)\\
%	\stackrel{\mathcal{Z}}{\leftrightarrow}\\
%\end{align*}

\bibliographystyle{plainnat}
\bibliography{lav}

\newpage
\begin{figure}[htp!]
  \begin{center}
    \includegraphics[scale=.6]{afr_hap_map.eps}
    \caption{Admixture fractions for 49 African American individuals in the HapMap 3 data. 
    Source population allele frequencies were estimated using 113 Yoruban and 111 European individuals.}
    \label{structure_plot}
  \end{center}
\end{figure}

\newpage
\begin{figure}[htp!]
  \begin{center}
    \includegraphics[scale=.6]{ensemble.eps}
    \caption{The admixture fractions of five replicate populations (each column) 5, 50, and 500 generations after an admixture pulse. As the admixture event grows more ancient, the variability within a replicate population decreases, but some variability is still maintained across the populations.}
    \label{ensemble}
  \end{center}
\end{figure}

\newpage
\begin{figure}[htp!]
  \begin{center}
    \includegraphics[scale=1.25]{verdu_comparison.eps}
    \caption{The variance predicted by \citet{verdu2011general} and equation \ref{var}, plotted on a logarithmic scale. The variance we predict (red) is always larger, but the two a very similar when $g$ is small.}
    \label{ek2_verdu}
  \end{center}
\end{figure}

\newpage
\begin{figure}[htp!]
  \begin{center}
    \includegraphics[scale=1.25]{rec_comparison.eps}
    \caption{The expected sample variance given by equation \ref{k2} plotted on a logarithmic scale, for a three different map functions. We used a map distance of $L=1$ Morgan and $N=10^4$. The Haldane map function ($1/2-e^{-2x}/2$) is in red, the Kosambi map function ($\tanh(2x)/2$) is in yellow, and the complete interence map function ($x$) is in blue. For all values of $g$, the expectations are ordered in the same order as the map functions, but the difference between the three disappears by $g=100$.}
    \label{ek2_rec}
  \end{center}
\end{figure}

\newpage
\begin{figure}[htp!]
  \begin{center}
    \includegraphics[scale=.6]{ci.eps}
    \caption{95\% confidence region for a model with constant admixture from generations $g_{start}$ to $g_{stop}$. The point estimate of $g_{start}=11$ and $g_{stop}=2$ generations ago is colored green.}
    \label{ci}
  \end{center}
\end{figure}

\newpage
\begin{figure}
    \begin{tabular}{ l | l | l | l }
    & Observed & Bootstrap & Corrected \\ \hline
$k_1$ & 0.777 & $-2.22\times 10^{-15}$& 0.777\\
$k_2$ & $9.00\times 10^{-3}$  &$2.59\times 10^{-4}$ & $8.75\times 10^{-3}$\\
$k_3$ & $2.98\times 10^{-4}$ & $1.60\times 10^{-5}$ & $2.82\times 10^{-4}$\\
$k_4$ & $-3.99\times10^{-5}$ & $-1.41\times 10^{-6}$ & $-3.85\times 10^{-5}$\\
    \end{tabular}
    \caption{$k$-statistics}
    \label{table:k_stats}
\end{figure}

%\newpage
%\section*{Math Appendix}
%For $i$ loci, we have a a state space of size $B_i$ which is the set of partions of $X=\{x_1,\dots,x_i\}$, $\mathcal{R}_X$.

%We will write the partitions as equivalence relations of $X$.
%These partitions form a lattice, with the partial order being refinement.
%For $A,B\in\mathcal{P}_i$, $A$ is a refinement of $B$ \textit{iid} $\forall a\in A\exists b\in B:a\subseteq b$.
%In this case, we write $A\leq B$.
%Moreover, this lattice is graded, by the function $r(A)\equiv |A|$.

%For every $P\in\mathcal{P}_i$ with $|P|=2$, we let $r(P)$ be the probability of generating that haplotype through recombination.
%To maintain consistency, we need for $Q\in \mathcal{P}_j$ with $j\leq i$,

%$$
%r(Q)=\sum_{P\in\mathcal{P}_i:\prec P}r(P)
%$$

%The recombination and drift matrices, indexed by partition, can now be defined element wise.
%\begin{align*}
%   L_{A,B} &= \left\{
%     \begin{array}{ll}
%       \frac{1}{N^{|A|}}\frac{N!}{(N-|B|)!} & : A\geq B\\
%       0 & : \text{else}
%     \end{array}
%   \right.\\
%   U_{A,B} &= \left\{
%     \begin{array}{ll}
%       1 & : A\geq B\\
%       0 & : \text{else}
%     \end{array}
%   \right.
%\end{align*}

\end{document}

\section*{Appendix}
The third $k$-statistic is an unbiased estimator of the third central moment and is given by
\begin{align*}
	k_3&=\frac{n^2}{(n-1)(n-2)}\sum_{i=1}^n\left(M_i-k_1\right)^3.
\end{align*}
Its expectation corresponds to the three-locus linkage disequilibrium as defined by (Bennett 195?) or (Slatkin 1972), 
which is a left eigenvalue of the three-locus drift and recombination matrices $\mathbf{L}_3$ and $\mathbf{U}_3$. 
Similar to the previous two, we get that
\begin{align*}
	\mathbb{E}(k_3)&=\mathbb{E}(M_1^3-3M_1^2M_2+2M_1M_2M_3)\\
	&=\frac{m(1-m)(2-m)}{L^3}\left(1-\frac{1}{2N}\right)^T\left(1-\frac{2}{2N}\right)^T\int_0^L\int_0^L\int_0^L[\ell\ell'\ell'']^Td\ell d\ell'd\ell''.
\end{align*}

The confidence intervals for these estimators depend on their variances. 
\begin{align*}
	\text{var}(k_1|\mathcal{M})&=\frac{\kappa_2}{n}\\
	\text{var}(k_2|\mathcal{M})&=\frac{\kappa_4}{n}+\frac{2\kappa^2_2}{n-1}\\
	\text{var}(k_3|\mathcal{M})&=\dots
\end{align*}
\begin{align*}
	\mathbb{E}(k_2^2)
	&=\frac{1}{(n-1)^2}\mathbb{E}\left(\sum_{i,j=1}^n M_i^2M_j^2-\frac{2}{n}\sum_{i,j,k=1}^n M_i^2M_jM_j+\frac{1}{n^2}\sum_{i,j,k,l=1}^n M_iM_jM_kM_l \right)\\
	&=\frac{1}{(n-1)^2}\bigg(n\mathbb{E}(M_1^4)+n(n-1)\mathbb{E}(M_1^2M_2^2)\\
	&\ \ \ -2\mathbb{E}(M_1^4)-4(n-1)\mathbb{E}(M_1^3M_2)-2(n-1)\mathbb{E}(M_1^2M_2^2)-2(n-1)(n-2)\mathbb{E}(M_1^2M_2M_3)\\
	&\ \ \ +\frac{1}{n}\mathbb{E}(M_1^4)+\frac{4(n-1)}{n}\mathbb{E}(M_1^3M_2)+\frac{3(n-1)}{n}\mathbb{E}(M_1^2M_2^2)+\frac{6(n-1)(n-2)}{n}\mathbb{E}(M_1^2M_2M_3)\\
	&\ \ \ +\frac{(n-1)(n-2)(n-3)}{n}\mathbb{E}(M_1M_2M_3M_4)\bigg)\\
	&=\frac{n+1}{n-1}\left(\mathbb{E}(M_1^2M_2^2)-2\mathbb{E}(M_1^2M_2M_3)+\mathbb{E}(M_1M_2M_3M_4)\right)\\
	&\ \ \ +\frac{1}{n}\left(\mathbb{E}(M_1^4)-4\mathbb{E}(M_1^3M_2)-3\mathbb{E}(M_1^2M_2^2)+12\mathbb{E}(M_1^2M_2M_3)-6\mathbb{E}(M_1M_2M_3M_4)\right)
\end{align*}

Central moments are not additive for $i\geq 4$, so this argument would not hold. 
\begin{align*}
	\mathbf{v}_{i(g)}&=\mathbf{L}_{i} \mathbf{D}_{i(g)}\mathbf{U}_{i}\mathbf{v}_{i(g)}
	+\sum_{j=1}^i s_{1,g}^j(1-s_{1,g})^{i-j}\mathbf{v}_{i-j(g-1)}
\end{align*}

\begin{align*}
	\mathbf{P}_{3(g)}&=\left( \begin{array}{c}
					 \mathbb{P}\{A_{1(g)}(\ell)=A_{1(g)}(\ell')=A_{1(g)}(\ell'')=1\}\\
					 \mathbb{P}\{A_{1(g)}(\ell)=A_{1(g)}(\ell')=A_{2(g)}(\ell'')=1\} 	\\
					 \mathbb{P}\{A_{1(g)}(\ell)=A_{2(g)}(\ell')=A_{2(g)}(\ell'')=1\} 	\\
					 \mathbb{P}\{A_{1(g)}(\ell)=A_{2(g)}(\ell')=A_{1(g)}(\ell'')=1\} 	\\
					 \mathbb{P}\{A_{1(g)}(\ell)=A_{2(g)}(\ell')=A_{3(g)}(\ell'')=1\}
				\end{array} \right)\\
	\mathbf{U}_3&=\left( \begin{array}{ccccc}
					 [\ell\ell'\ell''] 		& [\ell\ell'|\ell''] 	& [\ell|\ell'\ell''] & [\ell\ell''|\ell'] & [\ell|\ell'|\ell'']\\
					 0 	& [\ell\ell'] & 0 & 0 & [\ell|\ell']\\
					 0 	& 0 & [\ell'\ell''] & 0 & [\ell|\ell'']\\
					 0 	& 0 & 0 & [\ell\ell''] & [\ell'|\ell'']\\
					 0&0&0&0&1
				\end{array} \right)\\
	\mathbf{L}_3&=\frac{1}{N^2}\left( \begin{array}{ccccc}
					 N^2 		& 0 	& 0 & 0 & 0\\
					 N 	& N-1 & 0 & 0 & 0\\
					 N 	& 0 & N-1 & 0 & 0\\
					 N 	& 0 & 0 & N-1 & 0\\
					 1 & N-1 & N-1 & N-1 & (N-1)(N-2)
				\end{array} \right)\\
		\mathbf{D}_{3(g)}&=\left( \begin{array}{ccccc}
					 h_g		& 0 	& 0 & 0 & 0\\
					 0 	& h_g^2 & 0 & 0 & 0\\
					 0 	& 0 & h_{g}^2 & 0 & 0\\
					 0 	& 0 & 0 & h_{g}^2 & 0\\
					 0 & 0 & 0 & 0 & h_{g}^3
				\end{array} \right)
\end{align*}

### 2015-08-11 code/constant_hypothesis_tester.py

@@ -1,0 +1,57 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+
+statistics = ['k2','k3','k4']
+criticalValues = {s:{} for s in statistics}
+
+for fileName in os.listdir('simulations'):
+	if fileName[0] == 's':
+		T_start = int(fileName[6])
+		T_stop  = int(fileName[13:-4])
+		
+		#try:
+		data = np.loadtxt('simulations/'+fileName,skiprows=1)
+		#except ValueError,StopIteration:
+		#continue
+		
+		n  = len(data[0])
+		m1 = np.mean(data,axis=1)
+		m2 = np.mean((data-m1[:,None])**2,axis=1)
+		m3 = np.mean((data-m1[:,None])**3,axis=1)
+		m4 = np.mean((data-m1[:,None])**4,axis=1)
+
+		k2 = n*np.sort(m2)/(n-1)
+		k3 = n**2*np.sort(m3)/((n-1)*(n-2))
+		k4 = n**2*np.sort((n+1)*m4-3*(n-1)*m2**2)/((n-1)*(n-2)*(n-3))
+		
+		criticalValues['k2'][(T_start,T_stop)] = k2[[8,-8]]
+		criticalValues['k3'][(T_start,T_stop)] = k3[[8,-8]]
+		criticalValues['k4'][(T_start,T_stop)] = k4[[8,-8]]
+
+observed = {}
+observed['k2'] = 0.008636
+observed['k3'] = 0.000284
+observed['k4'] = -3.17950532473*10**-5
+
+
+image = np.zeros((17,17))
+
+for Ts in sorted(criticalValues['k2'].keys()):
+	rejected = False
+	for s in statistics:
+		if not criticalValues[s][Ts][0] < observed[s] < criticalValues[s][Ts][1]:
+			rejected = True
+	if not rejected:
+		try:
+			image[Ts] = 1
+		except IndexError:
+			pass
+	
+#image = image.transpose()
+
+plt.imshow(image,interpolation='nearest')
+plt.show()
+
+
+

### 2015-08-11 code/constant_migration.py

@@ -1,0 +1,31 @@
+import sys
+import itertools
+import multiprocessing as mp
+import wright_fisher_admixer as wf
+	
+k1 = 0.2264
+nSamples = 49
+
+def simulate(T):
+	T_start, T_stop = T
+	rate = 1-(1-k1)**(1/float(T_stop-T_start+1))
+
+	parameters	=	{}
+	for t in range(T_start,T_stop):
+		parameters[t]	=	([rate],[0])
+	parameters[T_stop] = ([rate,1-rate],[0,1])
+
+	sim	=	wf.Simulation(N=1000)
+
+	output = open('start=%i_stop=%i.txt' %T, 'w')
+	for i in xrange(10000):
+		d	= sim.drawDistribution(nSamples,30,parameters)
+		output.write(nSamples*'%f '%tuple(d[:,0]) + '\n')
+	#print	np.var(d[:,0],ddof=1),np.var(d[:,1],ddof=1),np.var(d[:,2],ddof=1)
+
+times = itertools.product(range(8),range(18))
+times = [(t_start, t_stop) for t_start, t_stop in times if t_start <= t_stop]
+
+pool = mp.Pool(processes=10)
+
+pool.map(simulate, times)

### 2015-08-11 code/mean_variance_plotter.py

@@ -1,0 +1,16 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+resamples = np.loadtxt('ASW_YRI_CEU_pruned_resample_B.txt')
+
+means = np.mean(resamples, axis=0)
+variances = np.var(resamples, axis=0, ddof=1)
+e3 = ((resamples - means)**3).sum(axis=0)
+
+x = means * (1-means)
+print np.polyfit(x, variances, 1)
+#plt.ylim([0,.0005])
+
+plt.scatter(x, variances)
+plt.show()
+

### 2015-08-11 code/pulse_hypothesis_tester.py

@@ -1,0 +1,29 @@
+import os
+import numpy as np
+
+results = {}
+resultsK3={}
+
+for fileName in os.listdir('simulations'):
+	if fileName[0] == 'T':
+		T = int(fileName[2:-4])
+		data = np.loadtxt('simulations/'+fileName)
+	
+		variances = np.sort(np.var(data,axis=1,ddof=1))
+		results[T] = variances[[12,-13]]
+
+		n  = len(data[0])
+		s1 = np.sum(data,axis=1)
+		s2 = np.sum(data**2,axis=1)
+		s3 = np.sum(data**3,axis=1)
+		k3 = np.sort(2*s1**3-3*n*s1*s2+n**2*s3)/(n*(n-1)*(n-2))
+		resultsK3[T] = k3[[12,-13]]
+
+
+observedVariance = 0.008636
+observedK3       = 0.000284
+
+for T in np.sort(results.keys()):
+	print T, 
+	print results[T][0] < observedVariance < results[T][1],
+	print resultsK3[T][0] < observedK3 < resultsK3[T][1]

### 2015-08-11 code/pulse_migration.py

@@ -1,0 +1,16 @@
+import sys
+import wright_fisher_admixer as wf
+	
+k1 = 0.2264
+T  = int(sys.argv[1])
+parameters	=	{}
+parameters[T]	=	([k1,1-k1],[0,1])
+
+
+nSamples = 49
+sim	=	wf.Simulation(N=1000)
+for i in range(1000):
+	d	= sim.drawDistribution(nSamples,30,parameters)
+	print nSamples*'%f '%tuple(d[:,0])
+	#print	np.var(d[:,0],ddof=1),np.var(d[:,1],ddof=1),np.var(d[:,2],ddof=1)
+

### 2015-08-11 code/wright_fisher_admixer.py

@@ -1,0 +1,174 @@
+import numpy as np
+from bisect import bisect
+class Simulation:
+	def __init__(self,N):
+		self.N	=	N
+	def drawDistribution(self,nSamples,R,admixtureParameters):	
+		population	=	Population(N=self.N)
+		population.makePedigree(nSamples,admixtureParameters)
+		population.inheritLocalAncestries(R)
+		distribution	=	np.zeros((nSamples,3))
+		for source in [0,1,2]:
+			for i,person in enumerate(population.generations[0].people.values()):
+				try:
+					distribution[i,source]	=	person.getAncestryFractions()[source]
+				except KeyError:
+					distribution[i,source]	=	0
+		return distribution
+class Population:
+	def __init__(self,N):
+		self.N						=	N
+	def makePedigree(self,nSamples,admixtureParameters):
+		currentGeneration	=	None
+		self.generations	=	[None for T in range(max(admixtureParameters)+1)]
+		for T in range(len(self.generations)):
+			try:
+				ms,sources	=	admixtureParameters[T]
+				currentGeneration	=	Generation(self,currentGeneration,nSamples,ms=ms,sources=sources)
+			except KeyError:
+				currentGeneration	=	Generation(self,currentGeneration,nSamples)
+			self.generations[T]	=	currentGeneration
+	def inheritLocalAncestries(self,R):
+		for generation in self.generations[::-1]:
+			generation.inheritLocalAncestries(R)
+class Generation:
+	def __init__(self,population,nextGeneration=None,nSamples=None,ms=None,sources=None):
+		if nextGeneration is None:
+			self.people	=	dict((i,Person(ms=ms,sources=sources)) for i in range(nSamples))
+		else:
+			leftParentIDs		=	[np.random.randint(population.N) for chromosome in nextGeneration]
+			rightParentIDs	=	[np.random.randint(population.N) for chromosome in nextGeneration]
+			parentIDs				=	set(leftParentIDs)|set(rightParentIDs)
+			
+			self.people	=	dict((ID,Person(ms=ms,sources=sources)) for ID in parentIDs)
+			nextGeneration.assignParents(self,leftParentIDs,rightParentIDs)
+	def __iter__(self):
+		return self
+	def next(self):
+		try: 
+			self.counter	+= 1
+		except AttributeError:
+			self.counter	=	0
[…]

### 2015-08-11 ci_plotter.py
Remove old files

@@ -1,73 +1,0 @@
-import os
-import numpy as np
-import matplotlib.pyplot as plt
-
-statistics = ['k2','k3','k4']
-criticalValues = {s:{} for s in statistics}
-
-for fileName in os.listdir('simulations'):
-	if fileName[0] == 's':
-		T_start = int(fileName[6])
-		T_stop  = int(fileName[13:-4])
-		
-		try:
-			data = np.loadtxt('simulations/'+fileName,skiprows=1)
-		except ValueError,StopIteration:
-			continue
-		
-		n  = len(data[0])
-		m1 = np.mean(data,axis=1)
-		m2 = np.mean((data-m1[:,None])**2,axis=1)
-		m3 = np.mean((data-m1[:,None])**3,axis=1)
-		m4 = np.mean((data-m1[:,None])**4,axis=1)
-
-		k2 = n*np.sort(m2)/(n-1)
-		k3 = n**2*np.sort(m3)/((n-1)*(n-2))
-		k4 = n**2*np.sort((n+1)*m4-3*(n-1)*m2**2)/((n-1)*(n-2)*(n-3))
-		
-		criticalValues['k2'][(T_start,T_stop)] = k2[[8,-8]]
-		criticalValues['k3'][(T_start,T_stop)] = k3[[8,-8]]
-		criticalValues['k4'][(T_start,T_stop)] = k4[[8,-8]]
-
-observed = {}
-print np.mean(m1)
-observed['k2'] = 0.008636
-observed['k3'] = 0.000284
-observed['k4'] = -3.17950532473*10**-5
-
-
-image = np.zeros((7,17))
-
-for Ts in sorted(criticalValues['k2'].keys()):
-	rejected = False
-	for s in statistics:
-		if not criticalValues[s][Ts][0] < observed[s] < criticalValues[s][Ts][1]:
-			rejected = True
-	if not rejected:
-		try:
-			image[Ts] = 1
-		except IndexError:
-			pass
-
-
-image[1,11:13] = 1
-image[2,7:16] = 1
-image[3,5:14] = 1
-image[4,4:9] = 1
-image[5,5:8] = 1
-
-image[2,11] = .5
-
-image = image.transpose()
-image = image[::-1,:]
-fig = plt.subplot(111)
-fig.imshow(image,interpolation='nearest')
-fig.set_yticks(range(17))
-fig.set_yticklabels(range(17)[::-1])
-fig.set_xlabel('g_stop')
-fig.set_ylabel('g_start')
-
-plt.show()
-
-
-

### 2015-08-11 constant_hypothesis_tester.py

@@ -1,57 +1,0 @@
-import os
-import numpy as np
-import matplotlib.pyplot as plt
-
-statistics = ['k2','k3','k4']
-criticalValues = {s:{} for s in statistics}
-
-for fileName in os.listdir('simulations'):
-	if fileName[0] == 's':
-		T_start = int(fileName[6])
-		T_stop  = int(fileName[13:-4])
-		
-		#try:
-		data = np.loadtxt('simulations/'+fileName,skiprows=1)
-		#except ValueError,StopIteration:
-		#continue
-		
-		n  = len(data[0])
-		m1 = np.mean(data,axis=1)
-		m2 = np.mean((data-m1[:,None])**2,axis=1)
-		m3 = np.mean((data-m1[:,None])**3,axis=1)
-		m4 = np.mean((data-m1[:,None])**4,axis=1)
-
-		k2 = n*np.sort(m2)/(n-1)
-		k3 = n**2*np.sort(m3)/((n-1)*(n-2))
-		k4 = n**2*np.sort((n+1)*m4-3*(n-1)*m2**2)/((n-1)*(n-2)*(n-3))
-		
-		criticalValues['k2'][(T_start,T_stop)] = k2[[8,-8]]
-		criticalValues['k3'][(T_start,T_stop)] = k3[[8,-8]]
-		criticalValues['k4'][(T_start,T_stop)] = k4[[8,-8]]
-
-observed = {}
-observed['k2'] = 0.008636
-observed['k3'] = 0.000284
-observed['k4'] = -3.17950532473*10**-5
-
-
-image = np.zeros((17,17))
-
-for Ts in sorted(criticalValues['k2'].keys()):
-	rejected = False
-	for s in statistics:
-		if not criticalValues[s][Ts][0] < observed[s] < criticalValues[s][Ts][1]:
-			rejected = True
-	if not rejected:
-		try:
-			image[Ts] = 1
-		except IndexError:
-			pass
-	
-#image = image.transpose()
-
-plt.imshow(image,interpolation='nearest')
-plt.show()
-
-
-

### 2015-08-11 constant_migration.py

@@ -1,31 +1,0 @@
-import sys
-import itertools
-import multiprocessing as mp
-import wright_fisher_admixer as wf
-	
-k1 = 0.2264
-nSamples = 49
-
-def simulate(T):
-	T_start, T_stop = T
-	rate = 1-(1-k1)**(1/float(T_stop-T_start+1))
-
-	parameters	=	{}
-	for t in range(T_start,T_stop):
-		parameters[t]	=	([rate],[0])
-	parameters[T_stop] = ([rate,1-rate],[0,1])
-
-	sim	=	wf.Simulation(N=1000)
-
-	output = open('start=%i_stop=%i.txt' %T, 'w')
-	for i in xrange(10000):
-		d	= sim.drawDistribution(nSamples,30,parameters)
-		output.write(nSamples*'%f '%tuple(d[:,0]) + '\n')
-	#print	np.var(d[:,0],ddof=1),np.var(d[:,1],ddof=1),np.var(d[:,2],ddof=1)
-
-times = itertools.product(range(8),range(18))
-times = [(t_start, t_stop) for t_start, t_stop in times if t_start <= t_stop]
-
-pool = mp.Pool(processes=10)
-
-pool.map(simulate, times)

### 2015-08-11 pulse_hypothesis_tester.py

@@ -1,29 +1,0 @@
-import os
-import numpy as np
-
-results = {}
-resultsK3={}
-
-for fileName in os.listdir('simulations'):
-	if fileName[0] == 'T':
-		T = int(fileName[2:-4])
-		data = np.loadtxt('simulations/'+fileName)
-	
-		variances = np.sort(np.var(data,axis=1,ddof=1))
-		results[T] = variances[[12,-13]]
-
-		n  = len(data[0])
-		s1 = np.sum(data,axis=1)
-		s2 = np.sum(data**2,axis=1)
-		s3 = np.sum(data**3,axis=1)
-		k3 = np.sort(2*s1**3-3*n*s1*s2+n**2*s3)/(n*(n-1)*(n-2))
-		resultsK3[T] = k3[[12,-13]]
-
-
-observedVariance = 0.008636
-observedK3       = 0.000284
-
-for T in np.sort(results.keys()):
-	print T, 
-	print results[T][0] < observedVariance < results[T][1],
-	print resultsK3[T][0] < observedK3 < resultsK3[T][1]

### 2015-08-11 pulse_migration.py

@@ -1,16 +1,0 @@
-import sys
-import wright_fisher_admixer as wf
-	
-k1 = 0.2264
-T  = int(sys.argv[1])
-parameters	=	{}
-parameters[T]	=	([k1,1-k1],[0,1])
-
-
-nSamples = 49
-sim	=	wf.Simulation(N=1000)
-for i in range(1000):
-	d	= sim.drawDistribution(nSamples,30,parameters)
-	print nSamples*'%f '%tuple(d[:,0])
-	#print	np.var(d[:,0],ddof=1),np.var(d[:,1],ddof=1),np.var(d[:,2],ddof=1)
-

### 2015-08-11 wright_fisher_admixer.py

@@ -1,174 +1,0 @@
-import numpy as np
-from bisect import bisect
-class Simulation:
-	def __init__(self,N):
-		self.N	=	N
-	def drawDistribution(self,nSamples,R,admixtureParameters):	
-		population	=	Population(N=self.N)
-		population.makePedigree(nSamples,admixtureParameters)
-		population.inheritLocalAncestries(R)
-		distribution	=	np.zeros((nSamples,3))
-		for source in [0,1,2]:
-			for i,person in enumerate(population.generations[0].people.values()):
-				try:
-					distribution[i,source]	=	person.getAncestryFractions()[source]
-				except KeyError:
-					distribution[i,source]	=	0
-		return distribution
-class Population:
-	def __init__(self,N):
-		self.N						=	N
-	def makePedigree(self,nSamples,admixtureParameters):
-		currentGeneration	=	None
-		self.generations	=	[None for T in range(max(admixtureParameters)+1)]
-		for T in range(len(self.generations)):
-			try:
-				ms,sources	=	admixtureParameters[T]
-				currentGeneration	=	Generation(self,currentGeneration,nSamples,ms=ms,sources=sources)
-			except KeyError:
-				currentGeneration	=	Generation(self,currentGeneration,nSamples)
-			self.generations[T]	=	currentGeneration
-	def inheritLocalAncestries(self,R):
-		for generation in self.generations[::-1]:
-			generation.inheritLocalAncestries(R)
-class Generation:
-	def __init__(self,population,nextGeneration=None,nSamples=None,ms=None,sources=None):
-		if nextGeneration is None:
-			self.people	=	dict((i,Person(ms=ms,sources=sources)) for i in range(nSamples))
-		else:
-			leftParentIDs		=	[np.random.randint(population.N) for chromosome in nextGeneration]
-			rightParentIDs	=	[np.random.randint(population.N) for chromosome in nextGeneration]
-			parentIDs				=	set(leftParentIDs)|set(rightParentIDs)
-			
-			self.people	=	dict((ID,Person(ms=ms,sources=sources)) for ID in parentIDs)
-			nextGeneration.assignParents(self,leftParentIDs,rightParentIDs)
-	def __iter__(self):
-		return self
-	def next(self):
-		try: 
-			self.counter	+= 1
-		except AttributeError:
-			self.counter	=	0
[…]

### 2015-08-11 paper/response.tex
Add response to reviewers and edits to manuscript

\documentclass[11pt]{amsart}
\usepackage{graphicx,amssymb,amsthm,setspace,fullpage,natbib}
\onehalfspacing

\title{Random Union of Gametes}
\author{Mason Liang}
\date{\today}

\begin{document}
\textit{Responses in italics}
\section{Reviewer 1}
The authors present a model for understanding how moments of admixture proportions in a 
sample of individuals from a population change over time. Unlike earlier methods that allow 
inference of admixture dates from ancestry segment lengths, their method should be applicable in 
situations where local ancestry estimates cannot be reliably obtained. 
In contrast to Verdu and Rosenberg 2011, who focus on the admixture fraction for a random 
individual at one locus, Liang and Nielsen study the mean admixture fraction of a haploid 
individual’s genome, which makes their results both more intuitive to me, but also directly 
interpretable in comparison to sample population data. [Specifically, I find their definition of H 
quite meaningful, whereas I could never figure out how $H_{1,g}$ in the earlier work could take on 
values different from 0, 1/2, or 1, since it describes the admixture proportions of one individual at 
exactly one locus, but that may be my own inflexible understanding.]\\
\textit{The random variable $H_{1,g}$ only takes on values from the set $\{0, 1/2, 1\}$, but its expectation can be equal to other values in the range $[0,1]$}\\
But furthermore they incorporate drift and the random recombination 
Aside from the numerous typos, in both text and equations, I have observed no substantial 
problems with the modeling, and find the applications to African American data well-chosen. It 
would have been neat to see the method applied to an older admixture event for which local 
ancestry estimates are not possible, but I understand it would be difficult to validate such an 
inference and older admixed datasets might not be readily available. Overall, I think this is an 
interesting and relevant work.\\ 
\textit{We used African American data because the }
\\
line 57 - delete second “model” \textit{This has been deleted.}\\
lines 116-117 equations: should the $M$ rather be $G$ in the equations? \textit{Yes, we have changed them to G's}\\
line 118 - ``factions" $\rightarrow$ ``fractions" \textit{We have added the r.}\\
line 120 - ``to the left" \textit{We have added the missing ``the''}\\
line 159 - is the location of the sites in genetic distance? \textit{, this clarification has been added to the manuscript.}\\
line 166 - delete second comma at end of equation \textit{We have deleted it.}\\
line 175 - ``interference" \textit{We have corrected the spelling.}\\
line 181 - ``this assumption" could you clarify which assumption (no crossover interference?) \textit{Yes, The line now reads ``However, none of the mathematical results of this paper will require this assumption of no crossover interference.''}\\
line 210 - I would have appreciated some explanation or definition of $\kappa_n$. \textbf{TBD}\\
line 229 - ``these are similar" \textit{We have added the missing ``are''.}\\
line 249 - ``becuase" \textit{We have corrected this.}\\
line 262 - do you define which population you infer to admix into the other? I assume you infer the migration rate of Europeans into an African population, but this is not clear from the text.
\textit{We have clarified the model ``We used a 3-parameter model of constant admixture in which the admixed population is entirely of African ancestry before generation $g_start$. For $g_{start}\leq g\leq g_{stop}$, the gene-flow probability from the European source population is $s_g=s$, with $s_g=0$ otherwise. ''}\\ 
line 263 - Did you explain which values you used in eqn 1? Ie, I assume you needed to choose $L$? Or if not, could you explain? \textit{We used $R=33$. This has been included in the text.}\\
line 291 - ``European and African" \text{This has been fixed.}\\
lines 293 and 295 – ``g\_stop" $\rightarrow$ ``$g_{stop}$" \textit{These have been fixed.}\\ 
line 294 - It is possible that these individuals may choose not to self-identify as African American, so an alternate explanation is that individuals with more recent admixture may not be in this self-identified population. \\
\textit{This is an explanation we had not considered. We have added it. ``Another possible explanation for this estimate is that individuals with more recent admixture tend to not self-identify, or were otherwise not included in the ASW population. The effects of this sampling bias would be the same as the decline in gene flow we observed.''}\\
Figure 2 - I would like this figure even more if it had estimates of $k_1$ and $k_2$ for each 
simulated population.\\
\textit{Figure 2 has been updated to include the sample variance for each replicate. We did not include the sample mean for simplicity and because it is more easily estimated by eye.}\\
Figure 3 - ``but the two are very similar"\\
\textit{We have added the missing ``re''.}

\section{Review 2}
This article proposes theoretical derivations for moments of the distribution of ancestry proportions in admixed populations. This is an interesting topic, and the authors propose a general recursive method to calculate high-order moments of the distribution that was not presented before. The manuscript takes as a starting point the model of Verdu and Rosenberg (VR) for the estimation of variance in admixed populations. VR does not take into account drift or the finiteness of the genome, and the present manuscript sets out to model these effects. The authors also distinguish between expectations within population (i.e., given a fixed population genealogy), and between populations (i.e., between individuals having independently sampled genealogies), and propose an inference scheme for structure data.

Even though this is an interesting manuscript, a few issues should be addressed.\\
\subsection{}
Gravel (2012) also discusses mean and variance in ancestry proportion in models with finite genome, drift,
and continuous migrations, although not all three jointly (pp 611-613 and 618-619). It also presents an
application to admixture timing in the ASW, also finding that variance in admixture proportion supports
more than a single pulse of admixture. The method was also used in Botigue et al 10.1073/pnas.1306223110.
This manuscript is more detailed than Gravel (2012) on this topic, but there is considerable overlap. At
the very least, the authors must acknowledge the overlap and compare the results in some detail.\\
\subsection{}
The equation at line 176.5 is comparable to the second term in equation (8) in Gravel (2012).
However, the first term of equation (8) (dominant in many practical cases) appears to be missing. Even
though the present manuscript takes into account correlations among individuals induced by the
genealogy, it does not seem to take into account the correlation among loci induced by genealogies.
If I understand things correctly, the authors start off with a "Diploid Wright-Fisher" model, but then
perform calculations according to a "Markovian Wright-Fisher" model (see Figure 2B in Gravel (2012)). If I
*don't* understand things correctly, I suppose that clarifications about exactly what expectations are
calculated would be useful (see below).

Fortunately, even if I am correct, I believe that the discrepancy can be corrected in the present models
using the law of total variance. At line 245.5, the authors do something similar, but condition the variance
over "population" genealogies. They could solve most of this problem by conditioning over "individual"
genealogies, as in Gravel (2012).

\subsection{}
The manuscript presents a moments method to estimate parameters directly from the distribution of
inferred admixture proportions. Even though distributions of ancestry proportions are indeed informative
of demography, its higher moments are also very sensitive to a diversity of effects that are not accounted
for here, such as assortative mating, population structure, nonrandom sampling, sex-biased admixture,
and correlated errors in ancestry assignments. If this method is to be considered for practical inference,
these caveats should be presented clearly to avoid mis-use. I have found practical application of variance
estimates to be perilous in most practical cases.

\subsection{}
By contrast, I suspect that drift and within-vs-cross-population effects will be small to negligibleexcept
for very small populations. This is not to say that it is not useful to model these, but it may be useful to
point this out. The authors show in figure 2 important differences in the mean ancestry proportion after
500 generations in an admixed population, but I found little detail as to how the simulation was
conducted. Gravel (2012) estimates the standard deviation in ancestry proportion across replicate
populations at 0.05 for a population of size 100, so I'm assuming that the simulation presented here uses
a population of similarly small size.

\subsection{}
Finally, if simulations have been performed, it would be a good idea to provide comparisons with the
theoretical results!

\subsection*{Minor points/clarifications}\ \\
Line 116.5: M is not defined. It may also be useful to spell out "the expectations are taken over ..."--I'm not
sure I understand exactly what was done. My understanding is that M is the population genealogy (which
yields a given distribution G), and that in the LHS terms have expectations taken over both populations and
individuals, whereas the RHS is taken over individuals in a fixed population.\\
Line 134-136
It could be good to clarify which expectations are used here as well, and what terms are being estimated.
For example, E[k1] corresponds to the LHS term at line 116.5 . However, E[k2] is an estimator for the RHS
term, since we are taking a sample mean and not the expected mean. It may be good to be explicit.\\
Line 144.5
It'd be useful to point the reader to the appendix for the definitions of the v's.\\
Line 249 "because"\\
\\
I don't always sign my reviews, but I gave so many hints at this point that I might as well--\\
Simon Gravel

\section{Reviewer 3}
Liang and Nielsen consider the joint distribution of admixture proportions $H_{i(g)}$, $i\in 1,\ldots, n$ of
n individuals sampled from an admixed population g generations after the start of admixture. Previous
work apparently explored the marginal distribution of the admixture proportion of a single individual
drawn from an admixed population. (This distribution was marginal both in the sense of applying,
marginally, to a single individual, and also being marginalized over all possible genealogies). By contrast,
Liang and Nielsen explore the joint distribution of H's conditional on the unknown genealogy connecting
the samples.

They derive expressions for the expected values of various k-statistics that can be computed from a
sample of H's, and then, by relating these expected values to admixture parameters in some special cases
they develop a method of moments estimator for these admixture parameters.

This is an interesting effort to relate admixture proportions, which are routinely estimated by programs
like structure (they are the individual q values), to the parameters of mechanistic models of population
admixture and could provide a valuable tool for increasing the interpretability of results from structure.

Below I provide a few comments that I hope will be helpful.

\subsection{} I found the author's implicit invocation of de Finetti's representation theorem and its extensions by
Hamos and Savage a little hard to wrap my head around here, primarily because it is not quite clear to me
what they are considering $\mathcal{G}$ to be. At lines 104 and 105 they seem to be making the case that
conditional on a certain realization of a random distribution the H's will be iid. But, then, later, they seem
to be equating $\mathcal{G}$ with a realization from the distribution of possible genealogies. I guess the
confusing part for me here is that if one is trying to compute the expected value of $k_2$, then presumably
you are conditioning on a certain realization of $\mathcal{G}$, and if that is the case and you are adhering to
the interpretation of $\mathcal{G}$ as being a realization from a random distribution a la the representation
theorem, then $H_{i(g)}$ and $H_{j(g)}$ should be conditionally independent and hence the terms involving the
expectations of their products would simplify quite a bit further than what we find above line 137. So, on
the one hand the authors are saying that the H's are conditionally iid, and on the other they are saying that
they are correlated and that is central to their derivation. It would be great if this could be clarified
somewhat.

\subsection{} One question I have as I read this is, "How far from sufficient are the H's?" The Hs themselves are going
to be functions of the underlying "cluster of origin" process estimate by, say, structure, when using the
linkage model. So, instead of boiling everything down to a single real number between 0 and 1 for each
individual (the H's), one might consider trying to do inference from, say, the distribution of the lengths of
tracts of common descent from different clusters. I suspect that might carry considerably more
information that is relevant to estimating admixture-model parameters than just the H's. How much
information is lost by just considering the H's? Some discussion of this would be appreciated.

\subsection{} Given 2, it seems that this method would be particularly appropriate in cases where a map is not
available --- i.e. in molecular ecology contexts studying non-model organisms. This being the case, it
would be good if the authors were able to address how robust (or not) the model might be to certain
departures from the sampling model and the mechanistic admixture model they employ. Of most concern
to me as regards the application of their method in molecular ecology contexts is the effect of population
structure within the admixed population. Population structure within the admixed population will clearly
alter the observed distribution of the H's. I think it would be good to know how the strength of an effect
like population structure compares to the strength of the admixture parameters in their model in altering
the distribution of the H's. I suspect that it could have a large effect that could complicate estimation of
the admixture parameters in populations with unknown spatial gradients of introgression, etc.

\subsection{} Echoing 3, a full set of simulations evaluating the power of the method across a range of scenarios (in
terms of number of markers and the differentiation between the two admixing populations) would be nice to
have, as well as assessments of the robustness (or not) of the model to population structure within the
admixed population. In some ways, it appears to me that the estimation of these sorts of admixture
parameters could be analogous to the estimation of K---the number of clusters---using the program
structure, an exercise that is fraught with peril and warrants caution because real populations seldom
behave like idealized ones. It would be nice if the caveats involved in this method could be stated up front.
(not that this would necessarily keep molecular ecologists from abusing it somewhat...the estimation of K
being a case in point).

\subsection{} The assumption that the estimation error of $H_{i(g)}$ is independent of $H_{i(g)}$ at line 208 seems like it
might not be the best assumption. A more reasonable assumption might be that the variance of $\epsilon_i$
is proportional to $H_i(1-H_i)$, as is common when dealing with proportions. Of course, this complicates the
calculations. Perhaps some type of variance-stabilzing transformation (like arc-sine square root) might be
useful.\\
\textit{TBD}

\subsection{} It would be nice to have a github repository with the all the scripts the authors used so that their
analyses are easily reproducible. In fact, it seems like some simple software (in R) for example,
implementing the authors' method is conspicuously absent. Many journals today require that data sets be
archived to insure reproducibility, and I think it is incumbent upon theoreticians, likewise, to offer clean
software or scripts to ensure reproducibility, and in some cases ease interpretation of methods.
\textit{The simulation code and scripts to reproduce the plots in the manuscript have been uploaded to github:  \texttt{https://github.com/z2trillion/local\_ancestry\_moments}.}

\subsection*{Minor points}
\subsubsection{} Equation above line 117. Should the $\mathcal{M}$ be $\mathcal{G}?$ If not, $\mathcal{M}$ needs definition. \textit{We have changed the $\mathcal{M}$ to a $\mathcal{G}$.}
\subsubsection{} I think that the second summation in the equation below line 136 should be over $i\neq j$, rather than
$i,j = 1$ up to n.
\textit{We have made this change.}
\subsubsection{} Line 148: Wait a moment (what a terrible pun!). Isn't $k_2$ estimating a central moment? Perhaps some more general discussion of $k$-statistics earlier in the paper would be good. 
\textit{The $k$-statistics are estimators for cumulants: $\mathbb{E}(k_i) = \kappa_i$ while the $h$-statistics are estimators for central moments: $\mathbb{E}(h_i) = \mu_i$. For $i=1,2,3$, the central moment and cumulant are the same, with $h_i$ and $k_i$ having the same formula, but for $i\geq 4$, the $i^\text{th}$ cumulant is not equal to the $i^\text{th}$ central moment.}
\subsubsection{} The manuscript would benefit from a few extra rounds of proofreading.
\end{document}

### 2015-08-11 paper/manuscript.tex
Add response to reviewers and edits to manuscript

@@
 \onehalfspacing
-\title{Estimating Admixture Parameters from Admixture Fractions}
+\title{Understanding Admixture Fractions}
 \author{Mason Liang and Rasmus Nielsen}
@@
 It is common in population genetic analyses to consider individuals as belonging fractionally to two or more discrete source populations.
-The proportion of an individual's genome that belongs to a population is called that individual's ``admixture fraction" or ``admixture proportion".
-Programs such as \verb|Structure| \citep{pritchard2000inference}, \verb|Eigenstrat|, \verb|Frappe| \citep{tang2005estimation}, or \verb|Admixture| \citep{alexander2009fast} can jointly estimate these admixture fractions for multiple individuals in a sample, along with the corresponding  allele frequencies in each of the source populations.
+The proportion of an individual's genome that belongs to a population is called that individual's `admixture fraction' or `admixture proportion'.
+Programs such as \verb|Structure| \citep{pritchard2000inference}, \verb|Eigenstrat| \citep{price2006principal}, \verb|Frappe| \citep{tang2005estimation}, or \verb|Admixture| \citep{alexander2009fast} can jointly estimate these admixture fractions for multiple individuals in a sample, along with the corresponding  allele frequencies in each of the source populations.
 These admixture fractions are often presented in a `structure plot,' an example of which is shown in Figure \ref{structure_plot}.
-We will henceforth refer to these methods as 'structure analyses'.
+We will henceforth refer to these methods as `structure analyses'.
 This approach has proven highly useful for understanding genetic relationships in many different species, e.g.
@@
 However, the admixture proportions themselves also contain information regarding admixture times.
-Right after an admixture event, the variance in admixture proportions within a population will be high, but will thereafter decrease, and will eventually (as we will show) converge to zero in the limit of large genomes.
+Following an admixture event, the variance in admixture proportions within a population will be high, but will thereafter decrease, and will eventually converge to zero in the limit of large genomes.
 The variance in admixture fractions among individuals contains substantial information about the time since admixture that can be used in addition to the tract length distribution.  
-It may in some cases be more robust than inferences based on tract lengths, because the length distribution of tracts is often difficult to infer, and is often not modeled accurately by the hidden Markov model (HMM) methods used to infer tract lengths \citep{liang2014lengths}.
+In some cases, this may be more robust than inferences based on tract lengths, because the length distribution of tracts is often difficult to infer, and is often not modeled accurately by the hidden Markov model (HMM) methods used to infer tract lengths \citep{liang2014lengths}.
 Even in cases where tract lengths can be accurately inferred, studies aimed at estimating admixture times should benefit from using both variance in admixture proportions among individuals and overall admixture tract lengths distributions.
@@
 
-We extend the model model in \citet{verdu2011general} to incorporate the effects of recombination and genetic drift by adding a a random union of gametes component.
+We extend the model in \citet{verdu2011general} to incorporate the effects of recombination and genetic drift by adding a a random union of zygotes component.
 Recombination is important because even if one half of a chromosome's ancestors are from the first source population, it is unlikely that exactly one half of that chromosome's genetic material is inherited from that population.
@@
 We use the same mechanistic admixture model of \citet{verdu2011general}, and will use its notation where possible.
-Finally, we use the random union of gametes model for genetic drift and recombination, and assume a sample size of $n$ chromosomes from a single population.
+Finally, we use the random union of zygotes model, with a diploid population size of $N$ ($2N$ chromosomes), for genetic drift and recombination, and assume a sample size of $n$ chromosomes from a single population.
 
-In the model of \citet{verdu2011general}, a hybrid population of $N$ diploid individuals forms in generation 1 from two isolated source populations.
+In this model, a hybrid population of $N$ diploid individuals forms in generation 1 from two previously isolated source populations.
 In this first generation, individuals in the hybrid population are from the first source population with probability $s_0$ or from the second source population with probability $1-s_0$.
-In subsequent generations, each individual in the hybrid population draws two recombined gametes at random from the previous generation.
-In generation $g+1$, each gamete is, independently, from the first source population with introgression probability $s_g$, or from the hybrid population with probability $1-s_g$.
-If the individual inherits both of its gametes from the hybrid population, the two gametes are from the same individual with probability $1/N$, and from different individuals with probability $1-1/N$.
+In generation $g+1$, each chromosome is, independently, from the first source population with introgression probability $s_g$, or from the hybrid population with probability $1-s_g$.
+Chromosomes inherited from the hybrid population are the product of the recombination of the two chromosomes of one individual (zygote), chosen uniformly at random.
+Finally, these $2N$ chromosomes are paired up to form the $N$ individuals in generation $g+1$.
 
-Mathematically, if the stochastic process $A(\ell)$ represents the local ancestry along a chromosome as a function of physical position ($\ell$):
+Finally, we let the stochastic process $A(\ell)$ represent the local ancestry along a chromosome as a function of $\ell$, the physical position:
 
@@
      \begin{array}{l}
-       0\ :\ \ell $ descended from first source population$\\
-       1\ :\ \ell $ descended from second source population$\\
+       0\ :\ \ell $ is descended from first source population$\\
+       1\ :\ \ell $ is descended from second source population$\\
      \end{array}
-   \right.,
+   \right. .
 $$
 
-then the fraction of the chromosome descended from the second source population is given by
+The fraction of the chromosome descended from the second source population is given by
 
@@
 We are interested in the joint distribution of these $n$ random variables.
-When $n=1$ and as $L\rightarrow\infty$, this is the admixture fraction considered in \citep{verdu2011general}.
+When $n=1$ and as $L\rightarrow\infty$, this is the admixture fraction considered by \citet{verdu2011general}.
 
 Because the $n$ chromosomes have possibly overlapping geneologies, the admixture fractions are not independent.
-However, the joint distribution of the admixture fractions does not depend on their ordering, so they are exchangeable (see appendix for mathematical definition). 
+However, the joint distribution of the admixture fractions does not depend on their ordering, so they are exchangeable. 
 As a result, they can be viewed as being identically and independently ($\textit{iid}$) drawn from a random distribution $\mathcal{G}$.
-This random distribution can be interpreted as a deterministic function of the random genealogy of the entire hybrid population up to $g$ generations in the past.
-When $g$ is small, the genealogies of the $n$ samples will be unlikely to differ from $n$ non-overlapping binary trees, so $\mathcal{M}$ will be approximately constant.
-If $g$ is large however, this will no longer hold.
+This random distribution can be interpreted as a function of the random genealogy of the entire hybrid population up to $g$ generations in the past.
+When $g$ is small, the genealogies of the $n$ samples will be unlikely to differ from $n$ non-overlapping binary trees, so $\mathcal{G}$ will be approximately constant.
+If $g$ is large however, these genealogies are likely to overlap, and this will no longer be true.
 
@@
 \begin{align*}
-\mathbb{E}(H_{1(g)})&\neq\mathbb{E}(H_{1(g)}|\mathcal{M})=\lim_{n\rightarrow\infty}\frac{1}{n}\sum_{i=1}^n H_{i(g)}\\
-\text{var}\left(H_{1(g)}\right)&\neq\text{var}(H_{1(g)}|\mathcal{M})=\lim_{n\rightarrow\infty}\frac{1}{n-1}\sum_{i=1}^n \left(H_{i(g)}-\frac{1}{n}\sum_{j=0}^n H_{j(g)}\right)^2.
+\mathbb{E}(H_{1(g)})&\neq\mathbb{E}(H_{1(g)}|\mathcal{G})=\lim_{n\rightarrow\infty}\frac{1}{n}\sum_{i=1}^n H_{i(g)}\\
+\text{var}\left(H_{1(g)}\right)&\neq\text{var}(H_{1(g)}|\mathcal{G})=\lim_{n\rightarrow\infty}\frac{1}{n-1}\sum_{i=1}^n \left(H_{i(g)}-\frac{1}{n}\sum_{j=0}^n H_{j(g)}\right)^2,
 \end{align*}
@@
 and similarly for higher-order moments.
-In other words, the moments of the admixture factions have two components: randomness due sampling of population history, and randomness due to the sampling of individuals.
-The expressions to left account for both, while the expressions to the right only account for the latter.
-Variances among individuals within a population will correspond to var$(H_{1(g)}|\mathcal{M})$ but var$(H_{1(g)})$ will tend to be larger than the expected sample variance calculated from multiple individuals sampled from the same population, and will rarely be useful for inference purposes.
+The moments of the admixture fractions have two components: randomness from sampling the population genealogy, and randomness from the sampling of chromosomes.
+The expressions to the left account for both, while the expressions to the right only account for the latter.
+Variances among individuals within one population correspond to var$(H_{1(g)}|\mathcal{G})$, while variances over replicate populations correspond to var$(H_{1(g)})$.
+This latter value will be larger than the expected sample variance calculated from multiple individuals sampled from the same population, and will rarely be useful for inference purposes.
 
@@
 
-We let $k_1$ be the sample mean:
+Let $k_1$ be the sample mean:
 
@@
 
-We let $k_2$ be the unbiased estimator of the sample variance:
+Similarly, let $k_2$ be the unbiased estimator of the sample variance:
 
@@
 	\mathbb{E}(k_2)
-	&=\frac{1}{n-1}\sum_{i=1}^n \mathbb{E}(H_{i,g}^2)-\frac{1}{n(n-1)}\sum_{i,j=1}^n \mathbb{E}(H_{i,g}H_{j,g})\\
+	&=\frac{1}{n-1}\sum_{i=1}^n \mathbb{E}(H_{i,g}^2)-\frac{1}{n(n-1)}\sum_{i\neq j} \mathbb{E}(H_{i,g}H_{j,g})\\
 	&=\mathbb{E}(H_{1,g}^2)-\mathbb{E}(H_{1,g}H_{2,g}).
@@
 
-These expectations can be written in terms of two-point correlation functions:
+These expectations can be written in terms of two-point correlation functions of $A$:
 
@@
 $$\mathbf{v}_{2(g)}=\left( \begin{array}{r}
-			\mathbb{P}\left\{A_{1,g}(\ell)=1, A_{1,g}(\ell')=1\right\} \\ \mathbb{P}\left\{A_{1,g}(\ell)=1, A_{2,g}(\ell')=1\right\}
+			\mathbb{P}\left\{A_{1(g)}(\ell)=1, A_{1(g)}(\ell')=1\right\} \\ \mathbb{P}\left\{A_{1(g)}(\ell)=1, A_{2(g)}(\ell')=1\right\}
 		\end{array} \right),$$
 
-we get that
+we find that
 
@@
 
-In general, the $i^\text{th}$ $k$-statistic is an unbiased estimator of the $i^\text{th}$ cumulant of $\mathcal{M}$, and its expectation can be written as an integral over $[0,L]^i $ of a linear combinations of $i$-point correlation functions.
+In general, the $i^\text{th}$ $k$-statistic is an unbiased estimator of the $i^\text{th}$ cumulant of $\mathcal{G}$, and its expectation can be written as an integral over $[0,L]^i $ of a linear combinations of $i$-point correlation functions.
 For example,
@@
 			1 & -1 &-1 &-1 &2	\end{array} \right)\mathbf{v}_{3(g)} d\ell d\ell'd\ell''\\
-	\mathbb{E}(k_4)&=\frac{1}{L^4}\int_0^L\int_0^L\int_0^L\int_0^L \Bigl(
+	\mathbb{E}(k_4)&=\frac{1}{L^4}\int_{[0,L]^4}\Bigl(
 			1 \  \underbrace{-1}_\text{4 times} \  \underbrace{-1}_\text{3 times} \  \underbrace{2}_\text{6 times} \ 6 \Bigr)\mathbf{v}_{4(g)} d\ell d\ell'd\ell''d\ell'''\\
@@
 
-Remarkably, the linear combinations required to compute the expectations of the $k$-statistics correspond exactly to the higher-order disequilibria as defined in [Bennett 1948?].
-Furthermore, if instead the we choose to compute the expectations of the $h$-statistics, which estimate the central moments, the linear combinations would correspond to the higher-order disequilibria as defined in [Slatkin 1972].
+Remarkably, the linear combinations required to compute the expectations of the $k$-statistics correspond exactly to the higher-order disequilibria as defined by \citet{bennett1952theory}.
+Furthermore, if instead the we choose to compute the expectations of the $h$-statistics, which estimate the central moments, the linear combinations would correspond to the higher-order disequilibria as defined by \citet{slatkin1972treating}.
 
@@
 $s_g=0$ for $g>0$.
-Using the random union of gametes model, we can compute $\mathbf{v}_{2(g)}$ in terms of the probabilities from the previous generation:
+Using the random union of zygotes model, we can compute $\mathbf{v}_{2(g)}$ in terms of the probabilities from the previous generation:
 
-If two sites at $\ell$ and $\ell'$ are on the same chromosome in generation $g$, then they were inherited from one gamete from one parent in the previous generation, while if they 	are on different haplotypes, then the probability that they are from gametes from one parent is $1/N$.
-On one gamete, the probability that these two sites are inherited from the same parental chromosome is $[\ell \ell']$ and the probability that they are instead inherited from different parental chromosomes is $[\ell|\ell']$.
+If two sites at genetic distances $\ell$ and $\ell'$ are on the same chromosome in generation $g+1$, then they were inherited from one chromosome from generation $g$ with probability $[\ell\ell']$ and from two chromosomes from generation $g$ with probability $[\ell|\ell']$.
+If they are on different chromosomes, then the probability that they are descended from one chromosome in generation $g$ is $\frac{1}{2N}[\ell\ell']$ and the probability that they are descended from two chromosomes is $\frac{1}{2N}[\ell|\ell']+\left(1-\frac{1}{2N}\right)$
 In matrix notation,
@@
 \begin{align*}
-\mathbf{v}_{2(g)}&=
-		\left(\mathbf{L}_2\mathbf{U}_2\right)^g
+\mathbf{v}_{2(g+1)}&=
+		\left(\mathbf{L}_2\mathbf{U}_2\right)
+		\mathbf{v}_{2(g)}
+		= \left(\mathbf{L}_2\mathbf{U}_2\right)^g
 		\mathbf{v}_{2(0)},
@@
 \begin{align*}
+	\mathbf{L}_2&=\left( \begin{array}{cc}
+					 1 & 0\\
+					 \frac{1}{2N} & 1-\frac{1}{2N}
+				\end{array} \right)\\
 		\mathbf{U}_2&=\left( \begin{array}{cc}
@@
 					 0 & 1
-				\end{array} \right)\\
-	\mathbf{L}_2&=\left( \begin{array}{cc}
-					 1 & 0\\
-					 1/N & 1-1/N
 				\end{array} \right).
@@
 This is the the same matrix equation (Wright 1933 and Hill and Robertson 1966) derived for the decay of two-locus linkage disequilibrium.
-The ``alleles" we consider are the local ancestry at $\ell$ and $\ell'$.
+The `alleles' we consider are the local ancestry at $\ell$ and $\ell'$.
 To the extent possible, our notation will follow (Hill 1974), whose results for measures of multi-locus linkage disequilibria we use.
-The matrices $\mathbf{L}_2$ and $\mathbf{U}_2$ share $(1\ -1)$ as a left-eigenvector, with corresponding eigenvalues $1-1/N$ and $[\ell\ell']$.
+The matrices $\mathbf{L}_2$ and $\mathbf{U}_2$ share $(1\ -1)$ as a left-eigenvector, with corresponding eigenvalues $1-\frac{1}{2N}$ and $[\ell\ell']$.
 As a result,
@@
 		\mathbf{v}_{2(0)} d\ell d\ell'\nonumber\\
-		&=\frac{1}{L^2}\left(1-\frac{1}{N}\right)^g\left( s_{1,0}-s_{1,0}^2 \right)\int_0^L\int_0^L [\ell\ell']^g d\ell d\ell'.
+		&=\frac{1}{L^2}\left(1-\frac{1}{2N}\right)^g\left( s_{0}-s_{0}^2 \right)\int_0^L\int_0^L [\ell\ell']^g d\ell d\ell'.
 	\label{k2_pulse}
@@
 \begin{align*}
-	\mathbb{E}(k_2)&=\frac{1}{L^2}\left(1-\frac{1}{N}\right)^g\left( s_{1,0}-s_{1,0}^2 \right)\int_0^L\int_0^L \left(\frac{1+\exp(-2|\ell-\ell'|)}{2}\right)^g d\ell d\ell',
+	\mathbb{E}(k_2)&=\frac{1}{L^2}\left(1-\frac{1}{2N}\right)^g\left( s_{0}-s_{0}^2 \right)\int_0^L\int_0^L \left(\frac{1+\exp(-2|\ell-\ell'|)}{2}\right)^g d\ell d\ell'\\
+	&=\frac{2}{L^2}\left(1-\frac{1}{2N}\right)^g\left( s_{0}-s_{0}^2 \right)\int_0^L (L-\ell)\left(\frac{1+\exp(-2\ell)}{2}\right)^g d\ell d\ell',
 \end{align*}
 
- while for a model of complete crossover inteference on a chromosome of length 1 Morgan, we can get a closed form solution:
+ while for a model of complete crossover interference on a chromosome of length 1 Morgan, we can get a closed form solution:
 
 \begin{align*}
-	\mathbb{E}(k_2)&=\left(1-\frac{1}{N}\right)^g\left( s_{1,0}-s_{1,0}^2 \right)\int_0^1\int_0^1 \left(1-|\ell-\ell'|\right)^g d\ell d\ell'\\
-	&=\left(1-\frac{1}{N}\right)^g \left(s_{1,0}-s_{1,0}^2\right) \frac{2}{2+g}.
+	\mathbb{E}(k_2)&=\left(1-\frac{1}{2N}\right)^g\left( s_{0}-s_{0}^2 \right)\int_0^1\int_0^1 \left(1-|\ell-\ell'|\right)^g d\ell d\ell'\\
+	&=\left(1-\frac{1}{2N}\right)^g \left(s_{0}-s_{0}^2\right) \frac{2}{2+g}.
 \end{align*}
@@
 For predicting the expected sample variance, the difference between these two models is not large, as shown in figure \ref{ek2_rec}.
-
 For the simulations and inference in this paper, we will ignore crossover interference, and use the Haldane map function.
-However, none of the mathematical results of this paper will require this assumption.
+However, none of the mathematical results of this paper will require this assumption of no crossover interference.
 
@@
 \begin{align*}
-	\mathbb{E}(k_3)&=\frac{s_0(1-s_0)(2-s_0)}{L^3}\left(1-\frac{1}{2N}\right)^T\left(1-\frac{2}{2N}\right)^T\int_0^L\int_0^L\int_0^L[\ell\ell'\ell'']^Td\ell d\ell'd\ell''.
+	\mathbb{E}(k_3)&=\frac{s_0(1-s_0)(2-s_0)}{L^3}\left(1-\frac{1}{2N}\right)^T\left(1-\frac{2}{2N}\right)^T\int_{[0,L]^3}[\ell\ell'\ell'']^Gd\ell d\ell'd\ell''.
 \end{align*}
 
-For $i\geq 4$, this is no longer true, but the results of [Hill 197?] can be used to compute $\textbf{v}_{i}$ without having to exponentiate the entire drift and recombination matrices.
-For example, for $k_4$, the drift and recombination matrices are $15\times 15$, but using the technique in [Hill 197?], we only need to exponentiate a $4\times 4$ matrix to compute $\mathbb{E}(k_4)$.
-%[Comment: I didn't understand the end of this sentence].
+For $i\geq 4$, this is no longer true, but the results of \citep{hill1974disequilibrium} can be used to compute $\textbf{v}_{i}(g)$ without having to exponentiate the entire drift and recombination matrices.
+For example, for $k_4$, the drift and recombination matrices are $15\times 15$, but using the technique in \citep{hill1974disequilibrium}, we only need to exponentiate a $4\times 4$ matrix to compute $\mathbb{E}(k_4)$.
 
@@
 
-where the diagonal matrix $\mathbf{D}_{i(g)}$ has entries giving the probabilities that all $|p|$ haplotypes in a correlation function are not inherited from migrant ancestors in the previous generation:
+where the diagonal matrix $\mathbf{D}_{i(g)}$ has entries giving the probabilities the set of chromosomes, $p$, in a correlation function are all from the hybrid population in the previous generation:
 
@@
 
-Note that as $s_{(g)}$ is a constant, then equation (\ref{with_migration}) is linear, and can be solved using a Laplace transform.
-
-%[insert stuff here]
+Note that if $s_{(g)}$ is fixed, then equation (\ref{with_migration}) is linear, and can be solved using a Laplace transform.
 
@@
 \begin{align*}
-	\mathbb{E}(k_n)&=\mathbb{E}\left(\kappa_n(H_{i(g)}+\epsilon_i|\mathcal{M})\right)\\
-	&=\mathbb{E}\left(\kappa_n(H_{i(g)}|\mathcal{M})\right)+\kappa_n(\epsilon_i).
+	\mathbb{E}(k_n)&=\mathbb{E}\left(\kappa_n(H_{i(g)}+\epsilon_i|\mathcal{G})\right)\\
+	&=\mathbb{E}\left(\kappa_n(H_{i(g)}|\mathcal{G})\right)+\kappa_n(\epsilon_i).
 \end{align*}
@@
 We have assumed that we have the ancestry fractions for each haplotype in the sample, but with genotyping data, we instead have their pairwise means: $(H_{1(g)}+H_{2(g)})/2\dots$.
-In turns out this is results in a decrease in the expectations of the $k$-statitics.
+This is results in a decrease in the expectations of the $k$-statitics.
 Conditional on the random distribution $\mathcal{G}$, $H_{1(g)},H_{2(g)},\dots$ are \textit{iid} drawn from $\mathcal{G}$.
@@
 		&=\mathbb{E}\left(\kappa_i\left(\frac{H_{1(g)}+H_{2(g)}}{2}\middle\vert\mathcal{G}\right)\right)\\
-		&=\mathbb{E}\left(\kappa_i\left(\frac{H_{1(g)}}{2}\middle\vert\mathcal{G}\right)+\kappa_i\left(\frac{H_{2(g)}}{2}\middle\vert\mathcal{M}\right)\right)\\
+		&=\mathbb{E}\left(\kappa_i\left(\frac{H_{1(g)}}{2}\middle\vert\mathcal{G}\right)+\kappa_i\left(\frac{H_{2(g)}}{2}\middle\vert\mathcal{G}\right)\right)\\
 		&=2^{-i+1}\mathbb{E}\left(\kappa_i\left(H_{1(g)}\middle\vert\mathcal{G}\right)\right)\\
@@
 The variance that \citep{verdu2011general} compute variance over all the replicate populations, while the variance we have computed in this paper is the expectation of the variance within a single population.
-When $g$ is small, these are not very different, but when $g$ is large, the variance within a population goes to zero, but the variance across the replicate populations does not.
+When $g$ is small, these are similar, but when $g$ is large, the variance within a population goes to zero, but the variance across the replicate populations does not.
 This effect is shown in Figure \ref{ek2_verdu}.
@@
 \begin{align}\nonumber
-	\text{var}(H_{1(g)})&=\mathbb{E}(H_{1,g}^2)-s_{1,0}^2\\
-		&=\frac{1}{L^2}\int_0^L\int_0^L\mathbb{P}\left\{A_{1,g}(\ell)=1, A_{1,g}(\ell')=1\right\} d\ell d\ell'-s_{1,0}^2\nonumber\\
-		&=\frac{1}{L^2}\left(s_{1,0}-s_{1,0}^2\right)\int_0^L\int_0^L1-\left(1-[\ell\ell']\right)\frac{1-[\ell\ell']^g\left(1-\frac{1}{N}\right)^g}{1-[\ell\ell']\left(1-\frac{1}{N}\right)} d\ell d\ell.
+	\text{var}(H_{1(g)})&=\mathbb{E}(H_{1,g}^2)-s_{0}^2\\
+		&=\frac{1}{L^2}\int_0^L\int_0^L\mathbb{P}\left\{A_{1,g}(\ell)=1, A_{1,g}(\ell')=1\right\} d\ell d\ell'-s_{0}^2\nonumber\\
+		&=\frac{1}{L^2}\left(s_{0}-s_{0}^2\right)\int_0^L\int_0^L1-\left(1-[\ell\ell']\right)\frac{1-[\ell\ell']^g\left(1-\frac{1}{2N}\right)^g}{1-[\ell\ell']\left(1-\frac{1}{2N}\right)} d\ell d\ell.
 		\label{var}
@@
 
-This variance and the expectation of the second $k$-statistic have the same limit as $N\rightarrow\infty$, but for finite $N$, the variance will always be larger.
+This variance and the expectation of the second $k$-statistic have the same limit as $N\rightarrow\infty$, but for finite $N$, the variance is larger.
 This is because
@@
 The first variance is small when $N$ is large, but is always non-negative.  
-This function is also plotted on figure \ref{ek2_verdu}.
+%This function is also plotted on figure \ref{ek2_verdu}. 
 The difference between this equation and equation \ref{k2} only becomes significant on a coalescent time scale.
-In the absence of genetic drift, the admixture fractions are approximately independent, becuase the samples do not share ancestors.
+In the absence of genetic drift, the admixture fractions are approximately independent, because the samples do not share ancestors.
 
-\subsection*{Application African American Date}
-We applied this method to a subset of the ASW, CEU, and YRI data from the HapMap 3 project [citation here].
+\subsection*{Application to African American Data}
+We applied this method to a subset of the ASW, CEU, and YRI data from the HapMap 3 project \citep{international2010integrating}.
 After excluding children from trios, there were the genotypes for 49 ASW, 113 YRI, and 112 CEU individuals. 
-We estimated the admixture fractions using the supervised learning mode of \verb|Admixture| [citation here], with the CEU and YRI individuals assigned to separate clusters. 
+We estimated the admixture fractions using the supervised learning mode of \verb|Admixture|, with the CEU and YRI individuals assigned to separate clusters. 
 The sampling distribution of the admixture fractions was estimated using the block bootstrap with $10^4$ replicates and 2678 blocks, giving a block size of approximately 10 CM.
@@
 
-We assumed a 3-parameter model of constant admixture.
-For $g_{start}\leq g\leq g_{stop}$, $s_g=s$ with $s_g=0$ elsewhere.
-By matching the block-bootstrap corrected $k_2$ and $k_3$ to the predictions of equation \ref{k2}, we obtained a point estimates of
+We used a 3-parameter model of constant admixture in which the admixed population is entirely of African ancestry before generation $g_start$.
+For $g_{start}\leq g\leq g_{stop}$, the gene-flow probability from the European source population is $s_g=s$, with $s_g=0$ otherwise.
+By matching the block-bootstrap corrected $k_2$ and $k_3$ to the predictions of equation \ref{k2} using a total recombination length of $R=33\text{ Morgans}$, we obtained point estimates of
 
@@
 \section*{Discussion}
-
 We have extended the mechanistic model of \citet{verdu2011general} to account for recombination and genetic drift.
@@
 This mechanistic model allows for a large number of parameters.
-For the purposes of inference, it seems that imposing constraints, i.e.
-a small number of pulses or constant admixture, will be needed to narrow the search space.
+For the purposes of inference, it seems that imposing constraints, i.e.\ a small number of pulses or constant admixture, will be needed to narrow the search space.
 
 In this paper, we have assumed that admixture only comes from one source population, this need not be the case.
-To account for admixture from multple source populations, equation \ref{k2} must be modified to account for the probability that haplotypes trace their descent to multiple source populations.
-Alogrithmically, this is feasible, but the notation is cumbersome.
+To account for admixture from multiple source populations, equation \ref{k2} must be modified to account for the probability that haplotypes trace their descent to multiple source populations.
+Algorithmically, this is feasible, but the notation is cumbersome.
 The resulting equations are given in the appendix, along with the equations for computing expectations of higher-order $k$-statistics.
 
-Applications of the method to African-American HapMap data provides estimates of the time since admixture between people of Europe and and African descent in America.
-Notice that the confidence set for the admixture parameters does not include values of $g_stop = 0$.
-We interpret this as evidence that admixture rates have declined the last two generations.
+Applications of the method to African-American HapMap data provides estimates of the time since admixture between people of European and African descent in America.
+Notice that the confidence set for the admixture parameters does not include values of $g_{stop} = 0$.
+We interpret this as evidence that admixture rates have declined the last few generations.
+The point estimate of time gene-flow stopped is $g_{stop} = 2$.
+This abrupt stop in gene-flow is a limitation of our model.
+A more realistic interpretation of this estimate would be that the rate of gene-flow has declined within the last 5 generations or so.
+Another possible explanation for this estimate is that individuals with more recent admixture tend to not self-identify, or were otherwise not included in the ASW population.
+The effects of this sampling bias would be the same as the decline in gene flow we observed.
 Also notice that admixture before 15 generations ago can be rejected.
@@
 Structure analyses have become one of the most commonly applied tools in population genomic analyses.
-The theory developed in this paper allows users of structure analyses to interpret their data in the context of an model of admixture between populations, and should find use in many studies aimed at understanding the history of populations.
-
-%\section*{Constant Admixture}
-%Suppose instead that every generation, a fraction, $m_1$, of the population is replaced by migrants from source population 1. 
-%There are now three parameters, $\mathbf{\Theta}=\{T,m_0,m_1\}$.
-%The recursive formulas become slightly more complex, but still solvable.
-%\begin{align*}
-%	\mathbb{P}\{A_{1,T}(\ell)=1\}&= m_1+(1-m_1)\mathbb{P}\{A_{1,T-1}(\ell)=1\}\\
-%		&=1-(1-m_1)^T\left(1-\mathbb{P}\{A_{1,T}(\ell)=1\}\right)\\
-%		&=1-(1-m_1)^T\left(1-m_0\right).
-%\end{align*}
-%\begin{align*}
-%	\mathbf{P}_T&=
-%		\mathbf{L}_2\mathbf{U}_2
-%		\left(\begin{array}{cc}1-m_1 & 0 \\0 & (1-m_1)^2\end{array}\right)
-%		\mathbf{P}_{T-1}+\left(\begin{array}{c}m_1 \\ m_1^2+2m_1\mathbb{P}\{A_{1,T-1}(\ell)=1\}\end{array}\right)\\
-%	\stackrel{\mathcal{Z}}{\leftrightarrow}\\
-%\end{align*}
+The theory developed in this paper allows users of structure analyses to interpret their data in the context of a model of admixture between populations, and should find use in many studies aimed at understanding the history of populations.
 
@@
   \begin{center}
-    \includegraphics[scale=.6]{ensemble.eps}
-    \caption{The admixture fractions of five replicate populations (each column) 5, 50, and 500 generations after an admixture pulse.
-As the admixture event grows more ancient, the variability within a replicate population decreases, but some variability is still maintained across the populations.}
+    \includegraphics[scale=.6]{ensemble_with_variance.eps}
+    \caption{The admixture fractions of five replicate populations (each row) 5, 50, and 500 generations after an admixture pulse.
+The number underneath each structure plot is $k_2$, the sample variance, for that replicate population.
+For more ancient admixture events, the variability within a replicate population decreases to 0, but some variability is still maintained across the populations.}
     \label{ensemble}
@@
     \caption{The variance predicted by \citet{verdu2011general} and equation \ref{var}, plotted on a logarithmic scale.
-The variance we predict (red) is always larger, but the two a very similar when $g$ is small.}
+The variance we predict (red) is always larger, but the two are very similar when $g$ is small.}
     \label{ek2_verdu}
@@
 
+\section*{Appendix}
+These are the matrices for computing $\mathbb{E}(k_3)$. The matrices for computing $\mathbb{E}(k_4)$ are $15\times  15$ and not given here, but can be found in \citep{hill1974disequilibrium}.
+
+\begin{align*}
+	\mathbf{v}_{3(g)}&=\left( \begin{array}{c}
+					 \mathbb{P}\{A_{1(g)}(\ell)=A_{1(g)}(\ell')=A_{1(g)}(\ell'')=1\}\\
+					 \mathbb{P}\{A_{1(g)}(\ell)=A_{1(g)}(\ell')=A_{2(g)}(\ell'')=1\} 	\\
+					 \mathbb{P}\{A_{1(g)}(\ell)=A_{2(g)}(\ell')=A_{2(g)}(\ell'')=1\} 	\\
+					 \mathbb{P}\{A_{1(g)}(\ell)=A_{2(g)}(\ell')=A_{1(g)}(\ell'')=1\} 	\\
+					 \mathbb{P}\{A_{1(g)}(\ell)=A_{2(g)}(\ell')=A_{3(g)}(\ell'')=1\}
+				\end{array} \right)\\
+	\mathbf{U}_3&=\left( \begin{array}{ccccc}
+					 [\ell\ell'\ell''] 		& [\ell\ell'|\ell''] 	& [\ell|\ell'\ell''] & [\ell\ell''|\ell'] & 0\\
+					 0 	& [\ell\ell'] & 0 & 0 & [\ell|\ell']\\
+					 0 	& 0 & [\ell'\ell''] & 0 & [\ell|\ell'']\\
+					 0 	& 0 & 0 & [\ell\ell''] & [\ell'|\ell'']\\
+					 0&0&0&0&1
+				\end{array} \right)\\
+	\mathbf{L}_3&=\frac{1}{4N^2}\left( \begin{array}{ccccc}
+					 4N^2 		& 0 	& 0 & 0 & 0\\
+					 2N 	& 2N-1 & 0 & 0 & 0\\
+					 2N 	& 0 & 2N-1 & 0 & 0\\
+					 2N 	& 0 & 0 & 2N-1 & 0\\
+					 1 & 2N-1 & 2N-1 & 2N-1 & (2N-1)(2N-2)
+				\end{array} \right)\\
+		\mathbf{D}_{3(g)}&=\left( \begin{array}{ccccc}
+					 1-s_g		& 0 	& 0 & 0 & 0\\
+					 0 	& (1-s_g)^2 & 0 & 0 & 0\\
+					 0 	& 0 & (1-s_{g})^2 & 0 & 0\\
+					 0 	& 0 & 0 & (1-s_{g})^2 & 0\\
+					 0 & 0 & 0 & 0 & (1-s_{g})^3
+				\end{array} \right)
+\end{align*}
+
+When there is migration from both source populations, the recursion relations for the $i$-point correlation functions will depend on $i-1$-point, $i-2$-point, $\dots$ correlations functions as well.
+As as example, consider the case of $\textbf{v}_{2(g)}$.
+Let the introgression probability from the second source population be given by $t_g$.
+The recursion equation for $\textbf{v}_{2(g)}$ now also depends on $\mathbf{v}_{1(g)}$.
+
+\begin{align*}
+\textbf{v}_{2(g+1)}&=\textbf{L}_2\left( \begin{array}{cc}
+					 1-s_g-t_g		& 0 \\
+					 0 	& (1-s_g-t_g)^2
+				\end{array} \right)\textbf{U}_2\textbf{v}_{2(g)}+ 
+				\left( \begin{array}{c}
+					 t_g \\
+					 t_g^2+2t_g\mathbb{P}\{A_{1(g)}(\ell)=1\}
+				\end{array} \right)\\
+				&=\textbf{L}_2\left( \begin{array}{cc}
+					 1-s_g-t_g		& 0 \\
+					 0 	& (1-s_g-t_g)^2
+				\end{array} \right)\textbf{U}_2\textbf{v}_{2(g)}+ 
+				\left( \begin{array}{c}
+					 t_g \\
+					 t_g^2+2t_g\mathbf{v}_{1(g)}
+				\end{array} \right).
+\end{align*}
+
+Similarly, the recursion equation for $\textbf{v}_{3(g)}$ depends on $\mathbf{v}_{2(g)}$ and $\mathbf{v}_{1(g)}$.
+
 \end{document}
@@
 \begin{align*}
-	\text{var}(k_1|\mathcal{M})&=\frac{\kappa_2}{n}\\
-	\text{var}(k_2|\mathcal{M})&=\frac{\kappa_4}{n}+\frac{2\kappa^2_2}{n-1}\\
-	\text{var}(k_3|\mathcal{M})&=\dots
+	\text{var}(k_1|\mathcal{G})&=\frac{\kappa_2}{n}\\
+	\text{var}(k_2|\mathcal{G})&=\frac{\kappa_4}{n}+\frac{2\kappa^2_2}{n-1}\\
+	\text{var}(k_3|\mathcal{G})&=\dots
 \end{align*}

### 2015-08-12 paper/response.tex
Add some more responses

@@
 values different from 0, 1/2, or 1, since it describes the admixture proportions of one individual at 
-exactly one locus, but that may be my own inflexible understanding.]\\
-\textit{The random variable $H_{1,g}$ only takes on values from the set $\{0, 1/2, 1\}$, but its expectation can be equal to other values in the range $[0,1]$}\\
+exactly one locus, but that may be my own inflexible understanding.]
 But furthermore they incorporate drift and the random recombination 
@@
 interesting and relevant work.\\ 
-\textit{We used African American data because the }
-\\
 line 57 - delete second “model” \textit{This has been deleted.}\\
@@
 
-\subsection{} I found the author's implicit invocation of de Finetti's representation theorem and its extensions by
-Hamos and Savage a little hard to wrap my head around here, primarily because it is not quite clear to me
-what they are considering $\mathcal{G}$ to be. At lines 104 and 105 they seem to be making the case that
-conditional on a certain realization of a random distribution the H's will be iid. But, then, later, they seem
-to be equating $\mathcal{G}$ with a realization from the distribution of possible genealogies. I guess the
-confusing part for me here is that if one is trying to compute the expected value of $k_2$, then presumably
-you are conditioning on a certain realization of $\mathcal{G}$, and if that is the case and you are adhering to
-the interpretation of $\mathcal{G}$ as being a realization from a random distribution a la the representation
-theorem, then $H_{i(g)}$ and $H_{j(g)}$ should be conditionally independent and hence the terms involving the
-expectations of their products would simplify quite a bit further than what we find above line 137. So, on
-the one hand the authors are saying that the H's are conditionally iid, and on the other they are saying that
-they are correlated and that is central to their derivation. It would be great if this could be clarified
-somewhat.
+\subsection{} I found the author's implicit invocation of de Finetti's representation theorem and its extensions by Hamos and Savage a little hard to wrap my head around here, primarily because it is not quite clear to me what they are considering $\mathcal{G}$ to be.
+At lines 104 and 105 they seem to be making the case that conditional on a certain realization of a random distribution the H's will be iid. 
+But, then, later, they seem to be equating $\mathcal{G}$ with a realization from the distribution of possible genealogies. 
+I guess the confusing part for me here is that if one is trying to compute the expected value of $k_2$, then presumably you are conditioning on a certain realization of $\mathcal{G}$, 
+and if that is the case and you are adhering to
+the interpretation of $\mathcal{G}$ as being a realization from a random distribution a la the representation theorem, 
+then $H_{i(g)}$ and $H_{j(g)}$ should be conditionally independent and hence the terms involving the
+expectations of their products would simplify quite a bit further than what we find above line 137. 
+So, on the one hand the authors are saying that the H's are conditionally iid, and on the other they are saying that they are correlated and that is central to their derivation. 
+It would be great if this could be clarified somewhat.\\
+\textit{
+	\indent
+	Some of this confusion arise from an error on our part in which we wrote $\mathcal{M}$ when we meant to write $\mathcal{G}$.
+	We apologize for this.
+	In this paper, we take $\mathcal{G}$ to be the genealogy of the admixed population.
+	Strictly speaking, this genealogy is more informative than the limiting ``empirical measure'' because many of the genealogical ancestors are not genetic ancestors, and so do not influence the observed admixture fractions.
+\\
+	\indent 
+	All of the expectations we compute in this paper (unless explicitly stated to be conditional), are over the distribution of $\mathcal{G}$.
+When marginalized over this random distribution, $H_{i(g)}$ and $H_{j(g)}$ are no longer independent, so computing $\mathbb{E}[k_i]$ becomes non-trivial.
+}
 
@@
 information that is relevant to estimating admixture-model parameters than just the H's. How much
-information is lost by just considering the H's? Some discussion of this would be appreciated.
+information is lost by just considering the H's? Some discussion of this would be appreciated.\\
+\textit{
+	The problem on inferring admixture times from the distribution of admixture tract lengths has been considered in papers such as Pool and Nielsen 2009, Gravel 2012, and Liang and Nielsen 2014. 
+	We see two key distinctions between using admixture fractions and admixture tracts.
+	The first is that admixture fractions may be available in situations where admixture tracts are not, such as when marker density is low or a recombination map is not available. 
+	Even when both are available, estimating admixture fractions is more robust than estimating admixture tract length distributions.  
+	The second is that finding the theoretical admixture tract length distribution is non-trivial. 
+	The mixture-of-exponentials approximation derived in Gravel 2012 is not always accurate.	
+}
 

### 2015-08-26 code/ensemble_maker.py
Add more code

@@ -1,0 +1,29 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+from admixture_ms import coalescent
+
+iterations = 5
+N = 500
+Ts = [5, 50, 500]
+r = 1.0
+m = .5
+n = 50
+c=coalescent(N=N, r=r)
+
+fig, axes = plt.subplots(nrows=3, ncols=iterations)
+for T, row in zip(Ts, axes):
+	for i, axis in enumerate(row):
+		sample=c.simulate_distribution(T=T, n=n, m=m)	
+		image = np.zeros((n, n))
+		for j, individual in enumerate(sample):
+			image[:round(n*individual), j] = 1
+		axis.imshow(image, interpolation='nearest')
+		k1	=	np.mean(sample)
+		k2	=	np.var(sample,ddof=1)
+		axis.set_xlabel('%.2e' %k2)
+		if i == 0:
+			axis.set_ylabel('T=%i' %T)
+		axis.set_xticks([])
+		axis.set_yticks([])
+plt.show()

### 2015-08-26 code/moments_sampler.py

@@ -1,0 +1,23 @@
+import numpy as np
+from admixture_ms import coalescent
+
+iterations	=	10**6
+N						=	1000
+T						=	20
+r						=	.01
+m						=	.3
+n						=	1000
+c=coalescent(N=N,r=r)
+
+for iteration in xrange(iterations):
+	sample=c.simulate_distribution(T=T,n=n,m=m)	
+	k1	=	np.mean(sample)
+	k2	=	np.var(sample,ddof=1)
+	k3	=	np.mean((sample-k1)**3)*n**2/(n-1)/(n-2)
+
+	m2	=	np.mean((sample-k1)**2)
+	m4	=	np.mean((sample-k1)**4)
+	k4	=	n**2*((n+1)*m4-3*(n-1)*m2**2)/(n-1)/(n-2)/(n-3)
+
+	print k1,k2,k3,k4
+

### 2015-08-26 paper/response.tex
Add some more responses to Gravel

@@
 genealogies, as in Gravel (2012).
+\textit{
+	We do not use a "Markovian Wright-Fisher" model in this paper.
+	The equation in this manuscript most comparable to equation (8) in Gravel (2012) is equation XX.
+	Equation YY in our manuscript gives the expectation of the second $k$-statistic,
+	in common usage, this would be the variance, but as we explain in lines .... in our case, it is not.
+	The variance of the admixture fractions (marginalizing over all genealogies) is computed in equation XX.
+	This does have two terms corresponding to the two in  equation (8) in Gravel (2012).
+	We have adopted the clearer terminology from that paper and now refer to the two as the "assortment variance" 
+	and "genealogy variance". 
+}
 
@@
 estimates to be perilous in most practical cases.
+\textit{
+	We have noted these caveats in 
+}
 
 \subsection{}
-By contrast, I suspect that drift and within-vs-cross-population effects will be small to negligibleexcept
+By contrast, I suspect that drift and within-vs-cross-population effects will be small to negligible except
 for very small populations. This is not to say that it is not useful to model these, but it may be useful to
@@
 a population of similarly small size.
+\textit{
+	We have included additional details about the simulations.
+	The simulations used a population size of 2N=1000, with m=0.5 and total recombination distance of 1 Morgan.
+	We have added these details to the paper. 
+	The results of our simulations are similar to those of Gravel (2012) and Rosenberg (????) when the admixture 
+	pulse is recent. 
+	As $T$ becomes larger, the effects of genetic drift become significant, and our simulation results diverge. 
+}
 
@@
 theoretical results!
+[Comment: I have simulations, I just need to make the plots.]
 

### 2015-08-26 paper/response.tex
Add yet more responses

@@
 \textit{
-	We have noted these caveats in 
+	We agree that these of deviations from the Wright-Fisher model and our model of admixture fraction
+	estimation  will impact our ability to make accurate inferences. 
+	Assortative mating, population structure, would tend to increase $\mathbb{E}[k_2]$, which would bias our 
+	estimates of the admixture time downwards. 
+	Correlated errors in ancestry assignment, on the other hand, would cause $k_2$ to be downwardly-biased,
+	which would make our estimate of the admixture time too big. 
 }

### 2015-12-06 paper/response.tex
Reflow reviewer 1 section

@@
 \section{Reviewer 1}
-The authors present a model for understanding how moments of admixture proportions in a 
-sample of individuals from a population change over time. Unlike earlier methods that allow 
-inference of admixture dates from ancestry segment lengths, their method should be applicable in 
-situations where local ancestry estimates cannot be reliably obtained. 
-In contrast to Verdu and Rosenberg 2011, who focus on the admixture fraction for a random 
-individual at one locus, Liang and Nielsen study the mean admixture fraction of a haploid 
-individual’s genome, which makes their results both more intuitive to me, but also directly 
-interpretable in comparison to sample population data. [Specifically, I find their definition of H 
-quite meaningful, whereas I could never figure out how $H_{1,g}$ in the earlier work could take on 
-values different from 0, 1/2, or 1, since it describes the admixture proportions of one individual at 
-exactly one locus, but that may be my own inflexible understanding.]
-But furthermore they incorporate drift and the random recombination 
-Aside from the numerous typos, in both text and equations, I have observed no substantial 
-problems with the modeling, and find the applications to African American data well-chosen. It 
-would have been neat to see the method applied to an older admixture event for which local 
-ancestry estimates are not possible, but I understand it would be difficult to validate such an 
-inference and older admixed datasets might not be readily available. Overall, I think this is an 
-interesting and relevant work.\\ 
-line 57 - delete second “model” \textit{This has been deleted.}\\
-lines 116-117 equations: should the $M$ rather be $G$ in the equations? \textit{Yes, we have changed them to G's}\\
-line 118 - ``factions" $\rightarrow$ ``fractions" \textit{We have added the r.}\\
-line 120 - ``to the left" \textit{We have added the missing ``the''}\\
-line 159 - is the location of the sites in genetic distance? \textit{, this clarification has been added to the manuscript.}\\
-line 166 - delete second comma at end of equation \textit{We have deleted it.}\\
-line 175 - ``interference" \textit{We have corrected the spelling.}\\
-line 181 - ``this assumption" could you clarify which assumption (no crossover interference?) \textit{Yes, The line now reads ``However, none of the mathematical results of this paper will require this assumption of no crossover interference.''}\\
-line 210 - I would have appreciated some explanation or definition of $\kappa_n$. \textbf{TBD}\\
-line 229 - ``these are similar" \textit{We have added the missing ``are''.}\\
-line 249 - ``becuase" \textit{We have corrected this.}\\
-line 262 - do you define which population you infer to admix into the other? I assume you infer the migration rate of Europeans into an African population, but this is not clear from the text.
-\textit{We have clarified the model ``We used a 3-parameter model of constant admixture in which the admixed population is entirely of African ancestry before generation $g_start$.
-For $g_{start}\leq g\leq g_{stop}$, the gene-flow probability from the European source population is $s_g=s$, with $s_g=0$ otherwise.
-''}\\
-line 263 - Did you explain which values you used in eqn 1? Ie, I assume you needed to choose $L$? Or if not, could you explain? \textit{We used $R=33$. This has been included in the text.}\\
-line 291 - ``European and African" \text{This has been fixed.}\\
-lines 293 and 295 – ``g\_stop" $\rightarrow$ ``$g_{stop}$" \textit{These have been fixed.}\\ 
-line 294 - It is possible that these individuals may choose not to self-identify as African American, so an alternate explanation is that individuals with more recent admixture may not be in this self-identified population.
-\\
-\textit{This is an explanation we had not considered.
-We have added it.
-``Another possible explanation for this estimate is that individuals with more recent admixture tend to not self-identify, or were otherwise not included in the ASW population.
-The effects of this sampling bias would be the same as the decline in gene flow we observed.''}\\
-Figure 2 - I would like this figure even more if it had estimates of $k_1$ and $k_2$ for each 
-simulated population.\\
-\textit{Figure 2 has been updated to include the sample variance for each replicate. We did not include the sample mean for simplicity and because it is more easily estimated by eye.}\\
+The authors present a model for understanding how moments of admixture
+proportions in a sample of individuals from a population change over time.
+Unlike earlier methods that allow inference of admixture dates from ancestry
+segment lengths, their method should be applicable in situations where local
+ancestry estimates cannot be reliably obtained. In contrast to Verdu and
+Rosenberg 2011, who focus on the admixture fraction for a random individual at
+one locus, Liang and Nielsen study the mean admixture fraction of a haploid
+individual’s genome, which makes their results both more intuitive to me, but
+also directly interpretable in comparison to sample population data.
+[Specifically, I find their definition of H quite meaningful, whereas I could
+never figure out how $H_{1,g}$ in the earlier work could take on values
+different from 0, 1/2, or 1, since it describes the admixture proportions of one
+individual at exactly one locus, but that may be my own inflexible
+understanding.] But furthermore they incorporate drift and the random
+recombination Aside from the numerous typos, in both text and equations, I have
+observed no substantial problems with the modeling, and find the applications to
+African American data well-chosen. It would have been neat to see the method
+applied to an older admixture event for which local ancestry estimates are not
+possible, but I understand it would be difficult to validate such an inference
+and older admixed datasets might not be readily available. Overall, I think this
+is an interesting and relevant work.\\
+line 57 - delete second “model”
+	\textit{This has been deleted.}\\
+lines 116-117 equations: should the $M$ rather be $G$ in the equations?
+	\textit{Yes, we have changed them to G's}\\
+line 118 - ``factions" $\rightarrow$ ``fractions"
+	\textit{We have added the r.}\\
+line 120 - ``to the left"
+	\textit{We have added the missing ``the''}\\
+line 159 - is the location of the sites in genetic distance?
+	\textit{, this clarification has been added to the manuscript.}\\
+line 166 - delete second comma at end of equation
+	\textit{We have deleted it.}\\
+line 175 - ``interference"
+	\textit{We have corrected the spelling.}\\
+line 181 - ``this assumption" could you clarify which assumption (no crossover
+interference?)
+	\textit{Yes, the line now reads ``However, none of the mathematical results of
+	this paper will require this assumption of no crossover interference.''}\\
+line 210 - I would have appreciated some explanation or definition of
+$\kappa_n$.
+	\textbf{We now define $\kappa_n$ on line LINE NUMBER GOES HERE.}\\
+line 229 - ``these are similar"
+	\textit{We have added the missing ``are''.}\\
+line 249 - ``becuase"
+	\textit{We have corrected this.}\\
+line 262 - do you define which population you infer to admix into the other? I
+assume you infer the migration rate of Europeans into an African population, but
+ this is not clear from the text.
+	\textit{We have clarified the model ``We used a 3-parameter model of constant
+	admixture in which the admixed population is entirely of African ancestry
+	before generation $g_start$. For $g_{start}\leq g\leq g_{stop}$, the gene-flow
+	probability from the European source population is $s_g=s$, with $s_g=0$
+	otherwise. ''}\\ line 263 - Did you explain which values you used in eqn 1?
+	Ie, I assume you needed to choose $L$? Or if not, could you explain?
+	\textit{We used $R=33$. This has been included in the text.}\\
+line 291 - ``European and African"
+	\text{This has been fixed.}\\
+lines 293 and 295 – ``g\_stop" $\rightarrow$ ``$g_{stop}$"
+	\textit{These have been fixed.}\\
+line 294 - It is possible that these individuals may choose not to self-identify
+as African American, so an alternate explanation is that individuals with more
+recent admixture may not be in this self-identified population. \\
+	\textit{This is an explanation we had not considered. We have added it.
+	``Another possible explanation for this estimate is that individuals with more
+	recent admixture tend to not self-identify, or were otherwise not included in
+	the ASW population. The effects of this sampling bias would be the same as the
+	decline in gene flow we observed.''}\\
+Figure 2 - I would like this figure even more if it had estimates of $k_1$ and
+$k_2$ for each simulated population.\\
+	\textit{Figure 2 has been updated to include the sample variance for each
+	replicate. For simplicity, we did not include the sample mean for simplicity
+	because it is more easily estimated by eye.}\\
 Figure 3 - ``but the two are very similar"\\

### 2015-12-12 paper/response.tex
Reflow text

@@
 
-\section{Review 2}
-This article proposes theoretical derivations for moments of the distribution of ancestry proportions in admixed populations.
-This is an interesting topic, and the authors propose a general recursive method to calculate high-order moments of the distribution that was not presented before.
-The manuscript takes as a starting point the model of Verdu and Rosenberg (VR) for the estimation of variance in admixed populations.
-VR does not take into account drift or the finiteness of the genome, and the present manuscript sets out to model these effects.
-The authors also distinguish between expectations within population (i.e., given a fixed population genealogy), and between populations (i.e., between individuals having independently sampled genealogies), and propose an inference scheme for structure data.
+\section{Review 2} This article proposes theoretical derivations for moments of
+the distribution of ancestry proportions in admixed populations. This is an
+interesting topic, and the authors propose a general recursive method to
+calculate high-order moments of the distribution that was not presented before.
+The manuscript takes as a starting point the model of Verdu and Rosenberg (VR)
+for the estimation of variance in admixed populations. VR does not take into
+account drift or the finiteness of the genome, and the present manuscript sets
+out to model these effects. The authors also distinguish between expectations
+within population (i.e., given a fixed population genealogy), and between
+populations (i.e., between individuals having independently sampled
+genealogies), and propose an inference scheme for structure data.
 
-Even though this is an interesting manuscript, a few issues should be addressed.\\
-\subsection{}
-Gravel (2012) also discusses mean and variance in ancestry proportion in models with finite genome, drift,
-and continuous migrations, although not all three jointly (pp 611-613 and 618-619). It also presents an
-application to admixture timing in the ASW, also finding that variance in admixture proportion supports
-more than a single pulse of admixture. The method was also used in Botigue et al 10.1073/pnas.1306223110.
-This manuscript is more detailed than Gravel (2012) on this topic, but there is considerable overlap. At
-the very least, the authors must acknowledge the overlap and compare the results in some detail.\\
-\subsection{}
-The equation at line 176.5 is comparable to the second term in equation (8) in Gravel (2012).
-However, the first term of equation (8) (dominant in many practical cases) appears to be missing. Even
-though the present manuscript takes into account correlations among individuals induced by the
-genealogy, it does not seem to take into account the correlation among loci induced by genealogies.
-If I understand things correctly, the authors start off with a "Diploid Wright-Fisher" model, but then
-perform calculations according to a "Markovian Wright-Fisher" model (see Figure 2B in Gravel (2012)). If I
-*don't* understand things correctly, I suppose that clarifications about exactly what expectations are
-calculated would be useful (see below).
+Even though this is an interesting manuscript, a few issues should be
+addressed.\\
 
-Fortunately, even if I am correct, I believe that the discrepancy can be corrected in the present models
-using the law of total variance. At line 245.5, the authors do something similar, but condition the variance
-over "population" genealogies. They could solve most of this problem by conditioning over "individual"
-genealogies, as in Gravel (2012).
-\textit{
-	We do not use a "Markovian Wright-Fisher" model in this paper.
-	The equation in this manuscript most comparable to equation (8) in Gravel (2012) is equation XX.
-	Equation YY in our manuscript gives the expectation of the second $k$-statistic,
-	in common usage, this would be the variance, but as we explain in lines .... in our case, it is not.
-	The variance of the admixture fractions (marginalizing over all genealogies) is computed in equation XX.
-	This does have two terms corresponding to the two in  equation (8) in Gravel (2012).
-	We have adopted the clearer terminology from that paper and now refer to the two as the "assortment variance"
-	and "genealogy variance".
-}
+\subsection{} Gravel (2012) also discusses mean and variance in ancestry
+proportion in models with finite genome, drift, and continuous migrations,
+although not all three jointly (pp 611-613 and 618-619). It also presents an
+application to admixture timing in the ASW, also finding that variance in
+admixture proportion supports more than a single pulse of admixture. The method
+was also used in Botigue et al 10.1073/pnas.1306223110. This manuscript is more
+detailed than Gravel (2012) on this topic, but there is considerable overlap. At
+the very least, the authors must acknowledge the overlap and compare the results
+in some detail.\\
+\subsection{} The equation at line 176.5 is comparable to the second term in
+equation (8) in Gravel (2012). However, the first term of equation (8) (dominant
+in many practical cases) appears to be missing. Even though the present
+manuscript takes into account correlations among individuals induced by the
+genealogy, it does not seem to take into account the correlation among loci
+induced by genealogies. If I understand things correctly, the authors start off
+with a "Diploid Wright-Fisher" model, but then perform calculations according to
+a "Markovian Wright-Fisher" model (see Figure 2B in Gravel (2012)). If I *don't*
+understand things correctly, I suppose that clarifications about exactly what
+expectations are calculated would be useful (see below).
 
-\subsection{}
-The manuscript presents a moments method to estimate parameters directly from the distribution of
-inferred admixture proportions. Even though distributions of ancestry proportions are indeed informative
-of demography, its higher moments are also very sensitive to a diversity of effects that are not accounted
-for here, such as assortative mating, population structure, nonrandom sampling, sex-biased admixture,
-and correlated errors in ancestry assignments. If this method is to be considered for practical inference,
-these caveats should be presented clearly to avoid mis-use. I have found practical application of variance
-estimates to be perilous in most practical cases.
-\textit{
-	We agree that these of deviations from the Wright-Fisher model and our model of admixture fraction
-	estimation  will impact our ability to make accurate inferences.
-	Assortative mating, population structure, would tend to increase $\mathbb{E}[k_2]$, which would bias our
-	estimates of the admixture time downwards.
-	Correlated errors in ancestry assignment, on the other hand, would cause $k_2$ to be downwardly-biased,
-	which would make our estimate of the admixture time too big.
-}
+Fortunately, even if I am correct, I believe that the discrepancy can be
+corrected in the present models using the law of total variance. At line 245.5,
+the authors do something similar, but condition the variance over "population"
+genealogies. They could solve most of this problem by conditioning over
+"individual" genealogies, as in Gravel (2012).
+	\textit{ We do not use a "Markovian Wright-Fisher" model in this paper. The
+	equation in this manuscript most comparable to equation (8) in Gravel (2012)
+	is equation XX. Equation YY in our manuscript gives the expectation of the
+	second $k$-statistic, in common usage, this would be the variance, but as we
+	explain in lines .... in our case, it is not. The variance of the admixture
+	fractions (marginalizing over all genealogies) is computed in equation XX.
+	This does have two terms corresponding to the two in  equation (8) in Gravel
+	(2012). We have adopted the clearer terminology from that paper and now refer
+	to the two as the "assortment variance" and "genealogy variance". }
 
-\subsection{}
-By contrast, I suspect that drift and within-vs-cross-population effects will be small to negligible except
-for very small populations. This is not to say that it is not useful to model these, but it may be useful to
-point this out. The authors show in figure 2 important differences in the mean ancestry proportion after
-500 generations in an admixed population, but I found little detail as to how the simulation was
-conducted. Gravel (2012) estimates the standard deviation in ancestry proportion across replicate
-populations at 0.05 for a population of size 100, so I'm assuming that the simulation presented here uses
-a population of similarly small size.
-\textit{
-	We have included additional details about the simulations.
-	The simulations used a population size of 2N=1000, with m=0.5 and total recombination distance of 1 Morgan.
-	We have added these details to the paper.
-	The results of our simulations are similar to those of Gravel (2012) and Rosenberg (????) when the admixture
-	pulse is recent.
-	As $T$ becomes larger, the effects of genetic drift become significant, and our simulation results diverge.
-}
+\subsection{} The manuscript presents a moments method to estimate parameters
+directly from the distribution of inferred admixture proportions. Even though
+distributions of ancestry proportions are indeed informative of demography, its
+higher moments are also very sensitive to a diversity of effects that are not
+accounted for here, such as assortative mating, population structure, nonrandom
+sampling, sex-biased admixture, and correlated errors in ancestry assignments.
+If this method is to be considered for practical inference, these caveats should
+be presented clearly to avoid mis-use. I have found practical application of
+variance estimates to be perilous in most practical cases.
+	\textit{ We agree that these of deviations from the Wright-Fisher model and
+	our model of admixture fraction estimation  will impact our ability to make
+	accurate inferences. Assortative mating, population structure, would tend to
+	increase $\mathbb{E}[k_2]$, which would bias our estimates of the admixture
+	time downwards. Correlated errors in ancestry assignment, on the other hand,
+	would cause $k_2$ to be downwardly-biased, which would make our estimate of
+	the admixture time too big. }
+
+\subsection{} By contrast, I suspect that drift and within-vs-cross-population
+effects will be small to negligible except for very small populations. This is
+not to say that it is not useful to model these, but it may be useful to point
+this out. The authors show in figure 2 important differences in the mean
+ancestry proportion after 500 generations in an admixed population, but I found
+little detail as to how the simulation was conducted. Gravel (2012) estimates
+the standard deviation in ancestry proportion across replicate populations at
+0.05 for a population of size 100, so I'm assuming that the simulation presented
+here uses a population of similarly small size.
+	\textit{ We have included additional details about the simulations. The
+	simulations used a population size of 2N=1000, with m=0.5 and total
+	recombination distance of 1 Morgan. We have added these details to the paper.
+	The results of our simulations are similar to those of Gravel (2012) and
+	Rosenberg (????) when the admixture pulse is recent. As $T$ becomes larger,
+	the effects of genetic drift become significant, and our simulation results
+	diverge. }
 
@@
 \section{Reviewer 3}
-Liang and Nielsen consider the joint distribution of admixture proportions $H_{i(g)}$, $i\in 1,\ldots, n$ of
-n individuals sampled from an admixed population g generations after the start of admixture. Previous
-work apparently explored the marginal distribution of the admixture proportion of a single individual
-drawn from an admixed population. (This distribution was marginal both in the sense of applying,
-marginally, to a single individual, and also being marginalized over all possible genealogies). By contrast,
-Liang and Nielsen explore the joint distribution of H's conditional on the unknown genealogy connecting
-the samples.
+Liang and Nielsen consider the joint distribution of admixture proportions
+$H_{i(g)}$, $i\in 1,\ldots, n$ of n individuals sampled from an admixed
+population g generations after the start of admixture. Previous work apparently
+explored the marginal distribution of the admixture proportion of a single
+individual drawn from an admixed population. (This distribution was marginal
+both in the sense of applying, marginally, to a single individual, and also
+being marginalized over all possible genealogies). By contrast, Liang and
+Nielsen explore the joint distribution of H's conditional on the unknown
+genealogy connecting the samples.
 
-They derive expressions for the expected values of various k-statistics that can be computed from a
-sample of H's, and then, by relating these expected values to admixture parameters in some special cases
-they develop a method of moments estimator for these admixture parameters.
+They derive expressions for the expected values of various k-statistics that can
+be computed from a sample of H's, and then, by relating these expected values to
+admixture parameters in some special cases they develop a method of moments
+estimator for these admixture parameters.
 
-This is an interesting effort to relate admixture proportions, which are routinely estimated by programs
-like structure (they are the individual q values), to the parameters of mechanistic models of population
-admixture and could provide a valuable tool for increasing the interpretability of results from structure.
+This is an interesting effort to relate admixture proportions, which are
+routinely estimated by programs like structure (they are the individual q
+values), to the parameters of mechanistic models of population admixture and
+could provide a valuable tool for increasing the interpretability of results
+from structure.
 
@@
 
-\subsection{} I found the author's implicit invocation of de Finetti's representation theorem and its extensions by Hamos and Savage a little hard to wrap my head around here, primarily because it is not quite clear to me what they are considering $\mathcal{G}$ to be.
-At lines 104 and 105 they seem to be making the case that conditional on a certain realization of a random distribution the H's will be iid.
-But, then, later, they seem to be equating $\mathcal{G}$ with a realization from the distribution of possible genealogies.
-I guess the confusing part for me here is that if one is trying to compute the expected value of $k_2$, then presumably you are conditioning on a certain realization of $\mathcal{G}$,
-and if that is the case and you are adhering to
-the interpretation of $\mathcal{G}$ as being a realization from a random distribution a la the representation theorem,
-then $H_{i(g)}$ and $H_{j(g)}$ should be conditionally independent and hence the terms involving the
-expectations of their products would simplify quite a bit further than what we find above line 137.
-So, on the one hand the authors are saying that the H's are conditionally iid, and on the other they are saying that they are correlated and that is central to their derivation.
-It would be great if this could be clarified somewhat.\\
+\subsection{}
+I found the author's implicit invocation of de Finetti's representation theorem
+and its extensions by Hamos and Savage a little hard to wrap my head around
+here, primarily because it is not quite clear to me what they are considering
+$\mathcal{G}$ to be. At lines 104 and 105 they seem to be making the case that
+conditional on a certain realization of a random distribution the H's will be
+iid. But, then, later, they seem to be equating $\mathcal{G}$ with a realization
+from the distribution of possible genealogies. I guess the confusing part for me
+here is that if one is trying to compute the expected value of $k_2$, then
+presumably you are conditioning on a certain realization of $\mathcal{G}$, and
+if that is the case and you are adhering to the interpretation of $\mathcal{G}$
+as being a realization from a random distribution a la the representation
+theorem, then $H_{i(g)}$ and $H_{j(g)}$ should be conditionally independent and
+hence the terms involving the expectations of their products would simplify
+quite a bit further than what we find above line 137. So, on the one hand the
+authors are saying that the H's are conditionally iid, and on the other they are
+saying that they are correlated and that is central to their derivation. It
+would be great if this could be clarified somewhat.\\
 \textit{
-	\indent
-	Some of this confusion arise from an error on our part in which we wrote $\mathcal{M}$ when we meant to write $\mathcal{G}$.
-	We apologize for this.
-	In this paper, we take $\mathcal{G}$ to be the genealogy of the admixed population.
-	Strictly speaking, this genealogy is more informative than the limiting ``empirical measure'' because many of the genealogical ancestors are not genetic ancestors, and so do not influence the observed admixture fractions.
-\\
-	\indent
-	All of the expectations we compute in this paper (unless explicitly stated to be conditional), are over the distribution of $\mathcal{G}$.
-When marginalized over this random distribution, $H_{i(g)}$ and $H_{j(g)}$ are no longer independent, so computing $\mathbb{E}[k_i]$ becomes non-trivial.
+	Some of this confusion may arise from an error on our part in which we wrote
+	$\mathcal{M}$ when we meant to write $\mathcal{G}$. We apologize for this. In
+	this paper, we take $\mathcal{G}$ to be the genealogy of the admixed
+	population. Strictly speaking, this genealogy is more informative than the
+	limiting ``empirical measure'' because many of the genealogical ancestors are
+	not genetic ancestors, and so do not influence the observed admixture
+	fractions. All of the expectations we compute in this paper (unless explicitly
+	stated to be conditional), are over the distribution of $\mathcal{G}$. When
+	marginalized over this random distribution, $H_{i(g)}$ and $H_{j(g)}$ are no
+	longer independent, so computing $\mathbb{E}[k_i]$ becomes non-trivial.
 }
 
-\subsection{} One question I have as I read this is, "How far from sufficient are the H's?" The Hs themselves are going
-to be functions of the underlying "cluster of origin" process estimate by, say, structure, when using the
-linkage model. So, instead of boiling everything down to a single real number between 0 and 1 for each
-individual (the H's), one might consider trying to do inference from, say, the distribution of the lengths of
-tracts of common descent from different clusters. I suspect that might carry considerably more
-information that is relevant to estimating admixture-model parameters than just the H's. How much
-information is lost by just considering the H's? Some discussion of this would be appreciated.\\
+\subsection{}
+One question I have as I read this is, "How far from sufficient are the H's?"
+The Hs themselves are going to be functions of the underlying "cluster of
+origin" process estimate by, say, structure, when using the linkage model. So,
+instead of boiling everything down to a single real number between 0 and 1 for
+each individual (the H's), one might consider trying to do inference from, say,
+the distribution of the lengths of tracts of common descent from different
+clusters. I suspect that might carry considerably more information that is
+relevant to estimating admixture-model parameters than just the H's. How much
+information is lost by just considering the H's? Some discussion of this would
+be appreciated.\\
 \textit{
-	The problem on inferring admixture times from the distribution of admixture tract lengths has been considered in papers such as Pool and Nielsen 2009, Gravel 2012, and Liang and Nielsen 2014.
-	We see two key distinctions between using admixture fractions and admixture tracts.
-	The first is that admixture fractions may be available in situations where admixture tracts are not, such as when marker density is low or a recombination map is not available.
-	Even when both are available, estimating admixture fractions is more robust than estimating admixture tract length distributions.
-	The second is that finding the theoretical admixture tract length distribution is non-trivial.
-	The mixture-of-exponentials approximation derived in Gravel 2012 is not always accurate.
+	The problem on inferring admixture times from the distribution of admixture
+	tract lengths has been considered in papers such as Pool and Nielsen 2009,
+	Gravel 2012, and Liang and Nielsen 2014. We see two key distinctions between
+	using admixture fractions and admixture tracts. The first is that admixture
+	fractions may be available in situations where admixture tracts are not, such
+	as when marker density is low or a recombination map is not available. Even
+	when both are available, estimating admixture fractions is more robust than
+	estimating admixture tract length distributions. The second is that finding
+	the theoretical admixture tract length distribution is non-trivial. The
+	mixture-of-exponentials approximation derived in Gravel 2012 is not always
+	accurate.
 }
 
-\subsection{} Given 2, it seems that this method would be particularly appropriate in cases where a map is not
-available --- i.e. in molecular ecology contexts studying non-model organisms. This being the case, it
-would be good if the authors were able to address how robust (or not) the model might be to certain
-departures from the sampling model and the mechanistic admixture model they employ. Of most concern
-to me as regards the application of their method in molecular ecology contexts is the effect of population
-structure within the admixed population. Population structure within the admixed population will clearly
-alter the observed distribution of the H's. I think it would be good to know how the strength of an effect
-like population structure compares to the strength of the admixture parameters in their model in altering
-the distribution of the H's. I suspect that it could have a large effect that could complicate estimation of
-the admixture parameters in populations with unknown spatial gradients of introgression, etc.
+\subsection{} Given 2, it seems that this method would be particularly
+appropriate in cases where a map is not available --- i.e. in molecular ecology
+contexts studying non-model organisms. This being the case, it would be good if
+the authors were able to address how robust (or not) the model might be to
+certain departures from the sampling model and the mechanistic admixture model
+they employ. Of most concern to me as regards the application of their method in
+molecular ecology contexts is the effect of population structure within the
+admixed population. Population structure within the admixed population will
+clearly alter the observed distribution of the H's. I think it would be good to
+know how the strength of an effect like population structure compares to the
+strength of the admixture parameters in their model in altering the distribution
+of the H's. I suspect that it could have a large effect that could complicate
+estimation of the admixture parameters in populations with unknown spatial
+gradients of introgression, etc.\\
+\textit{
+}
 
-\subsection{} Echoing 3, a full set of simulations evaluating the power of the method across a range of scenarios (in
-terms of number of markers and the differentiation between the two admixing populations) would be nice to
-have, as well as assessments of the robustness (or not) of the model to population structure within the
-admixed population. In some ways, it appears to me that the estimation of these sorts of admixture
-parameters could be analogous to the estimation of K---the number of clusters---using the program
-structure, an exercise that is fraught with peril and warrants caution because real populations seldom
-behave like idealized ones. It would be nice if the caveats involved in this method could be stated up front.
-(not that this would necessarily keep molecular ecologists from abusing it somewhat...the estimation of K
-being a case in point).
+\subsection{}
+Echoing 3, a full set of simulations evaluating the power of the
+method across a range of scenarios (in terms of number of markers and the
+differentiation between the two admixing populations) would be nice to have, as
+well as assessments of the robustness (or not) of the model to population
+structure within the admixed population. In some ways, it appears to me that the
+estimation of these sorts of admixture parameters could be analogous to the
+estimation of K---the number of clusters---using the program structure, an
+exercise that is fraught with peril and warrants caution because real
+populations seldom behave like idealized ones. It would be nice if the caveats
+involved in this method could be stated up front. (not that this would
+necessarily keep molecular ecologists from abusing it somewhat...the estimation
+of K being a case in point).\\
+\textit{
+}
 
-\subsection{} The assumption that the estimation error of $H_{i(g)}$ is independent of $H_{i(g)}$ at line 208 seems like it
-might not be the best assumption. A more reasonable assumption might be that the variance of $\epsilon_i$
-is proportional to $H_i(1-H_i)$, as is common when dealing with proportions. Of course, this complicates the
-calculations. Perhaps some type of variance-stabilzing transformation (like arc-sine square root) might be
+\subsection{}
+The assumption that the estimation error of $H_{i(g)}$ is
+independent of $H_{i(g)}$ at line 208 seems like it might not be the best
+assumption. A more reasonable assumption might be that the variance of
+$\epsilon_i$ is proportional to $H_i(1-H_i)$, as is common when dealing with
+proportions. Of course, this complicates the calculations. Perhaps some type of
+variance-stabilzing transformation (like arc-sine square root) might be
 useful.\\
-\textit{TBD}
+\textit{
+	TBD
+}
 
-\subsection{} It would be nice to have a github repository with the all the scripts the authors used so that their
-analyses are easily reproducible. In fact, it seems like some simple software (in R) for example,
-implementing the authors' method is conspicuously absent. Many journals today require that data sets be
-archived to insure reproducibility, and I think it is incumbent upon theoreticians, likewise, to offer clean
-software or scripts to ensure reproducibility, and in some cases ease interpretation of methods.
-\textit{The simulation code and scripts to reproduce the plots in the manuscript have been uploaded to github:  \texttt{https://github.com/z2trillion/local\_ancestry\_moments}.}
+\subsection{}
+It would be nice to have a github repository with the all the
+scripts the authors used so that their analyses are easily reproducible. In
+fact, it seems like some simple software (in R) for example, implementing the
+authors' method is conspicuously absent. Many journals today require that data
+sets be archived to insure reproducibility, and I think it is incumbent upon
+theoreticians, likewise, to offer clean software or scripts to ensure
+reproducibility, and in some cases ease interpretation of methods.\\
+\textit{The
+simulation code and scripts to reproduce the plots in the manuscript have been
+uploaded to github:
+\texttt{https://github.com/z2trillion/local\_ancestry\_moments}.}
 
 \subsection*{Minor points}
-\subsubsection{} Equation above line 117. Should the $\mathcal{M}$ be $\mathcal{G}?$ If not, $\mathcal{M}$ needs definition. \textit{We have changed the $\mathcal{M}$ to a $\mathcal{G}$.}
-\subsubsection{} I think that the second summation in the equation below line 136 should be over $i\neq j$, rather than
-$i,j = 1$ up to n.
-\textit{We have made this change.}
-\subsubsection{} Line 148: Wait a moment (what a terrible pun!). Isn't $k_2$ estimating a central moment? Perhaps some more general discussion of $k$-statistics earlier in the paper would be good.
-\textit{The $k$-statistics are estimators for cumulants: $\mathbb{E}(k_i) = \kappa_i$ while the $h$-statistics are estimators for central moments: $\mathbb{E}(h_i) = \mu_i$.
-For $i=1,2,3$, the central moment and cumulant are the same, with $h_i$ and $k_i$ having the same formula, but for $i\geq 4$, the $i^\text{th}$ cumulant is not equal to the $i^\text{th}$ central moment.}
-\subsubsection{} The manuscript would benefit from a few extra rounds of proofreading.
+\subsubsection{}
+Equation above line 117. Should the
+$\mathcal{M}$ be $\mathcal{G}?$ If not, $\mathcal{M}$ needs definition.\\
+\textit{
+	We have changed the $\mathcal{M}$ to a $\mathcal{G}$.
+}
+
+\subsubsection{}
+I think that the second summation in the equation below line 136 should be over
+$i\neq j$, rather than $i,j = 1$ up to n.\\
+\textit{
+	Our original version was wrong. We have made this change.
+}
+
+\subsubsection{}
+Line 148: Wait a moment (what a terrible pun!). Isn't $k_2$ estimating a central
+moment? Perhaps some more general discussion of $k$-statistics earlier in the
+paper would be good. \\
+\textit{
+	The $k$-statistics are estimators for cumulants: $\mathbb{E}(k_i) = \kappa_i$
+	while the $h$-statistics are estimators for central moments: $\mathbb{E}(h_i)
+	= \mu_i$. For $i=1,2,3$, the central moment and cumulant are the same, with
+	$h_i$ and $k_i$ having the same formula, but for $i\geq 4$, the $i^\text{th}$
+	cumulant is not equal to the $i^\text{th}$ central moment. In the revised
+	paper, we explain this difference in lines TODO(mason).
+}
+
+\subsubsection{} The manuscript would benefit from a few extra rounds of
+proofreading.\\
+\textit{
+	We have gone through a couple more times.
+}
 \end{document}
